### 2017-03-10 QPQ9/qpq9_draft.tex
Remove .idea files

%\documentclass[12pt,preprint]{aastex}
\documentclass[iop]{emulateapj}
\usepackage{natbib,amsmath,graphicx,longtable,amssymb,latexsym,epsf}
\begin{document}

\title{Quasars Probing Quasars IX. The Kinematics of the Circumgalactic Medium Surrounding 
$z\sim2$ Quasars [v1.4]}

\author{Marie Wingyee Lau\altaffilmark{1,2}, J. Xavier Prochaska\altaffilmark{2}, 
Joseph F. Hennawi\altaffilmark{3}
}
\altaffiltext{1}{Email: lwymarie@ucolick.org}
\altaffiltext{2}{Department of Astronomy and Astrophysics, UCO/Lick Observatory, University of 
California, 1156 High Street, Santa Cruz, CA 95064}
\altaffiltext{3}{Department of Physics, Broida Hall, University of California, Santa Barbara, 
CA 93106-9530}

\begin{abstract}
We examine the kinematics of the gas in the environments of galaxies hosting quasars at $z\sim2$. 
We employ 112 projected quasar pairs to study the circumgalactic gas of the foreground quasars in 
absorption. The sample selects foreground quasars with precise redshift measurements, using 
emission-lines with precision $\lesssim300\,{\rm km\,s^{-1}}$ and average offsets from the 
systemic redshift $\lesssim|100\,{\rm km\,s^{-1}}|$. We stack the background quasar spectra at the 
foreground quasar's systemic redshift to study the mean absorption in \ion{C}{2}, \ion{C}{4}, 
and \ion{Mg}{2}. We find that the mean absorptions exhibit large velocity widths 
$\sigma_v>300\,{\rm km\,s^{-1}}$. We show that, gas in gravitational motion, Hubble flow, or 
galactic-scale outflows are all consistent with the observed widths. Furthermore, we find that the 
mean absorptions are asymmetric about the systemic redshift. The mean absorption centroids exhibit 
small redshift relative to the systemic $\delta v\approx+200\,{\rm km\,s^{-1}}$, with large 
intrinsic scatter in the centroid velocities of the individual absorption systems. We show 
that the observed offsets may be produced by anisotropic or intermittent ionizing radiation from 
the foreground quasars. 
\end{abstract}

\keywords{glaxies: clusters: intracluster medium -- galaxies: formation -- galaxies: halos -- 
intergalactic medium -- quasars: absorption lines -- quasars: general}

\section{Introduction}
\label{sec:introduction}

Galaxy formation and evolution are driven by the flows of gas into and out of their interstellar 
medium. Current theories demand that star-forming galaxies maintain these flows. Gas accretes, 
cools, and adds to the fuel supply, while star formation feedback heats gas, blows it out of 
galaxies, and regulates star formation \citep[for a review see][]{SomervilleDave15}.

Direct observations of galactic flows are difficult to acquire. Detecting the presence of 
the gas is itself challenging. Either the gas mass is too small, or the gas density is too low for 
the detection of line-emission, e.g.\ 21\,cm, Ly$\alpha$, or H$\alpha$ from \ion{H}{1}. Resolving 
the kinematics and establishing the mass flux pose an even greater challenge. These challenges are 
accentuated for distant, young galaxies, where flows of gas are predicted to prevail 
\citep{Keres+09,Fumagalli+11}. Therefore, with rare exceptions, 
\citep[e.g.][]{Cantalupo+14,Hennawi+15}, the community has relied on absorption-line spectroscopy 
to detect and characterize the gas surrounding galaxies 
\citep[e.g.][]{BergeronBoisse91,Steidel+10,Prochaska+11,Tumlinson+13}. 
In a previous paper of the Qussars Probing Quasars series \citep[][, hereafter QPQ8]{QPQ8}, we 
measured the velocity field for \ion{C}{2}\,1334 and \ion{C}{4}\,1548, finding that the 
circumgalactic medium frequently exhibits large velocity widths that are offset from the systemic 
redshift. The offsets, $\delta v$, are often positive, with the sign convention that positive 
velocities indicate a redshift from the systemic. From a sample of 7 \ion{C}{2} systems and 10 
\ion{C}{4} systems, we measured the velocity interval that encompasses 90\% of the total optical 
depth, $\Delta v_{90}$, and the $1\sigma$ dispersion relative to the profile centroid, $\sigma_v$. 
The median $\Delta v_{90}$ is $555\,{\rm km\,s^{-1}}$ for \ion{C}{2}\,1334 and is 
$342\,{\rm km\,s^{-1}}$ for \ion{C}{4}\,1548. The median $\sigma_v$ is $249\,{\rm km\,s^{-1}}$ for 
\ion{C}{2}\,1334 and is $495\,{\rm km\,s^{-1}}$ for \ion{C}{4}\,1548. These velocity fields exceed 
all previous measurements from galaxies and/or absorption systems at any epoch.

With absorption-line spectroscopy of background sightlines, other researchers also have had 
success in characterizing the flows of gas around galaxies. \cite{Rakic+12} found a net 
large-scale inflow around star-forming galaxies, or a Kaiser effect for gas on 
1\textendash2\,Mpc scales. \cite{Johnson+15} studied the CGM surrounding $z\sim1$ quasars, and 
found large peculiar motions in the gas exceeding the expected virial velocity, which they 
interpret as quasar-driven outflows. However, the velocity spreads were not well quantified. Other 
existing studies that found large velocity spreads are limited to single sightlines with most of 
the gas within several hundred ${\rm km\,s^{-1}}$ \citep[e.g., ][]{Tripp+11}, or with gas tracing 
a higher ionization state than the QPQ absorption systems \citep[e.g., ][]{Churchill+12}, or where 
the average velocity spread is smaller than that measured in QPQ8 
\citep[e.g., ][]{Gauthier13,Muzahid+15,Zahedy+16}.

A significant limitation of absorption-line analysis of transverse sightlines, especially 
regarding galactic-scale flows, is the inherent symmetry of the experiment. One generally lacks 
any constraint on the distance of the gas along the sightline. Positive or negative velocities 
with respect to the galaxy may be interpreted as gas flowing either toward or away from the 
system. ``Down-the-barrel'' observations break this symmetry, and have generally provided evidence 
for flows away from galaxies \citep{Rupke+05,Martin05,Weiner+09,Rubin+14}. However, these data are 
frequently at low spectral resolution which limits one's sensitivity to inflowing gas.

In this paper (hereafter QPQ9), we examine the flows of gas in the environments of massive 
galaxies hosting quasars. Our approach leverages a large dataset of quasar pairs 
\citep[][, hereafter QPQ1]{QPQ1} to use the standard techniques of absorption-line spectroscopy 
with background quasars. These quasar pairs 
have angular separations that correspond to less than 300\,kpc projected separation at the 
foreground quasar's redshift. Our previous publications from these quasar pairs have established 
that these galaxies are surrounded by a massive, cool, and enriched CGM 
\citep[QPQ5, QPQ6, QPQ7:][]{QPQ5,QPQ6,QPQ7}. We have collected a sample of 112 background 
spectra that are paired with foreground quasars with precisely measured redshifts. 
Among the sightlines in the QPQ9 sample, 13 have spectral resolution $R>5000$ from echellette or 
echelle observations, and have been analyzed separately in QPQ8. Our primary 
scientific interests are twofold: (i) search for signatures of galactic-scale outflows from the 
central galaxy, presumably driven by recent star formation and/or active galactic nuclei feedback; 
(ii) characterize the dynamics of the gas around these massive systems. We further describe 
an aspect of this experiment that offers a unique opportunity to study galactic-scale flows: we 
argue that, the anisotropic or intermittent radiation from the foreground quasars may break the 
symmetry in the velocity field of circumgalactic absorbers. If the ionizing radiation field is 
asymmetric, the absorbers may also distribute asymmetrically. Alternatively, finite quasar 
lifetime will result in different radiation fields impinged on the gas closer to versus 
further away from the background quasar, due to different light travel times. 
\cite{KirkmanTytler08} reported an asymmetry in \ion{H}{1} absorption on scales larger than the 
CGM, and gave similar arguments on anisotropy or intermittence.

%The formation and evolution of stars are dictated
%by the flow of gas and metals/dust.  [This transport
%varies directly with the stellar stage
%and a lesser extent stellar mass.]
%The accretion onto a protostellar  core eventually
%initiates nuclear fusion.  The main sequence is characterized
%by gentler stellar winds, with the most massive
%stars shedding a cumulative mass of XX\%.
%Stellar deaths, meanwhile, are frequently accompanied by terrific
%mass loss, occasionally punctuated by a highly energetic,
%supernova phase.
%By analogy, galaxy formation and evolution are driven by the
%flows into/out of their interstellar medium.
%In contrast to most stars, however, current theories
%predict (even demand) that the star-forming galaxies
%maintin these flows. [add]
%Akin to stars, direct observations of galactic flows are
%difficult to acquire. Even detecting the gas is challenging:
%the gas mass is either too small or the gas density too low
%for the [detection] of line-emission (e.g. 21cm, lya,
%or H$\alpha$ from Hydrogen).  Resolving the kinematics
%and establish the mass-flux represent a far greater challenge.
%These challenges are accentuated for distant, young galaxies
%where flows are predicted to prevail \citep{keres,fumagalli}.
%Therefore, with rare exceptions \citep[e.g.][]{slug,jackpot},
%the community has relied on absorption-line spectroscopy
%to detect and characterize the gas surrounding galaxies
%\citep[e.g.][]{bergeron,steidel10,pro11,tumlinson13}.
%With this experiment, researcher have had recent success
%in characterizing the large-scale flows around star-forming
%galaxies to reveal a net inflow \cite{rakic13} and
%provide unique constraints on the mass of the underlying
%potential well \cite{rakic14}.
%A significant limitation of standard absorption-line
%analysis, especially regarding sutyding galactic flows,
%is the inherent symmetry of the experiment.
%Because one generally lacks any constraint on the distance
%of the gas along the sightline,
%postive or negative velocities with respect to the
%galaxy may be interpreted as gas flowing either towards
%or away from the system.
%So-called `down-the-barrel' observations break this
%symmetry, and have generally provided evidence for flows
%away from galaxies \citep{rupke,martin,weiner,rubin}.
%However, these data are frequently at low spectral resolution
%which limits one's sensitivity to inflowing gas.
%In this paper, we examine the flows of gas in the environments
%of massive galaxies hosting quasars.  Our approach leverages
%a large dataset of quasar pairs \citep{hennawi}
%to use the standard technique of absorption-line spectroscopy
%with background quasars. [word]
%These quasar pairs have angular
%separations that correspond to less than 300\,kpc physical
%separation at the f/g quasar redshift.
%Our previous publications from these quasar pairs
%have established that these galaxies are surrounded
%by a massive, cool, and enriched circumgalactic medium
%\citep{QPQ1,QPQ5,QPQ6,QPQ7}.
%We have collected a sample of ntot~spectra passing within
%300\,kpc from a f/g quasar with a precisely measured
%redshift.  Our primary scientific interests are twofold:
%(i) search for signatures of galactic-scale outflows from the
%central galaxy, presumably driven by recent star-formation
%and/or AGN feedback;
%(ii) characterize the dynamics of galactic flows around these
%massive system.  We further descrbie an aspect of this
%experiment that offers a unique opportunity to study galactic
%flows. [detail??]

We adopt a $\Lambda$CDM cosmology with $\Omega_M=0.26, \Omega_\Lambda=0.74$, and 
$H_0=70\,{\rm km\,s^{-1}\,Mpc^{-1}}$. Distances are proper unless otherwise stated. When referring 
to comoving distances we include explicitly an $h^{-1}$ term and follow modern convention of 
scaling to a Hubble constant of $70\,{\rm km\,s^{-1}\,Mpc^{-1}}$. 
%All equivalent width measurements are presented in the rest-frame, unless otherwise specified.

\section{The Experiment}
\label{sec:experiment}

\input{Tables/tab_sample.tex}
%\input{tab_sample.tex}

\begin{figure}
\includegraphics[width=3in]{Figures/fig_experiment.pdf}
%\includegraphics[width=3in]{f01.pdf}
\caption{These panels summarize properties of the QPQ9 dataset. The QPQ survey selects 
quasar pairs of $R_\perp < 300$\,kpc and $z_{\rm fg}>1.6$. Assuming that the foreground quasars 
emit isotropically and at a distance equal to the impact parameter, the enhancement in the UV 
flux relative to the extragalactic UV background, $g_{\rm UV}$, can be estimated. Large 
symbols correspond to foreground quasars with the most precise redshift measurement from 
[\ion{O}{3}]\,5007, while small symbols correspond to $z_{\rm fg}$ 
measurements from \ion{Mg}{2}\,2800, H$\alpha$, or H$\beta$ emission. The top panel shows 
quasar pairs with coverage of \ion{C}{2}\,1334 at $z_{\rm fg}$ in the background quasar spectra. 
The middle panel shows pairs with coverage of \ion{C}{4}\,1548. The bottom panel shows 
pairs with coverage of \ion{Mg}{2}\,2796. 
}
\label{fig:experiment}
\end{figure}

The goal of our experiment is to measure the average velocity fields of the absorption from C$^+$, 
C$^{3+}$, and Mg$^+$ ions associated with the CGM of the galaxies hosting $z\sim 2$ quasars.

From our QPQ survey\footnote{http://www.qpqsurvey.org}, we analyze a subset of systems that 
pass within transverse separation $R_\perp<300$\,kpc from a foreground quasar with 
$z_{\rm fg} > 1.6$. We restrict the sample to foreground quasars with redshift measured from 
\ion{Mg}{2}\,2800, [\ion{O}{3}]\,5007, or H$\alpha$ emission, giving a precision of 
$300\,{\rm km\,s^{-1}}$ or better and an average offset from the systemic redshift of 
$|100\,{\rm km\,s^{-1}}|$ or less. According to \cite{Shen+16}, the [\ion{O}{3}] emission-line 
redshifts have the smallest scatter (intrinsic scatter and measurement error combined) of 
$68\,{\rm km\,s^{-1}}$ about the systemic redshift, and we analyze the sub-sample with 
[\ion{O}{3}] redshifts separately. The [\ion{O}{3}] line has an average blueshift of 
$48\,{\rm km\,s^{-1}}$ about the systemic redshift, which has been added when we compute the 
redshift of the line. The scatter and average offset of [\ion{O}{3}] redshifts reported by 
\cite{Shen+16} is consistent with the numbers reported by \cite{Boroson05} using a larger but 
lower redshift sample. Systemic redshifts measured from \ion{Mg}{2} have a precision of 
$226\,{\rm km\,s^{-1}}$ according to \cite{Shen+16}, and we have taken into account their 
reported median blueshift of $57\,{\rm km\,s^{-1}}$ of \ion{Mg}{2} from the systemic. We note that 
\cite{Richards+02} reported a median redshift of $97\,{\rm km\,s^{-1}}$ of \ion{Mg}{2} from 
[\ion{O}{3}] using a larger but lower redshift sample. In QPQ8, we quantified 
the precision of H$\alpha$ to be $300\,{\rm km\,s^{-1}}$ and the median offset from the systemic 
redshift is close to zero, consistent with the velocity shifts measured by \cite{Shen+11}. 
Although H$\beta$ is a narrow emission-line, we do not consider its redshift sufficiently reliable 
for use as systemic redshift. H$\beta$ redshifts have a large scatter about the systemic 
$\approx400\,{\rm km\,s^{-1}}$, and a large average offset about the systemic 
$\approx100\,{\rm km\,s^{-1}}$ \citep[][QPQ8]{Shen+16}. 
We further add to the QPQ dataset with quasar pairs selected from the public dataset of 
igmspec\footnote{https://github.com/specdb/igmspec}, which includes the spectra from the quasar 
catalogs based upon the Sloan Digital Sky Survey Seventh Data Release \citep{Schneider+10} and the 
Twelfth Data Release \citep{Paris+17}. We only select pairs with $z_{\rm fg}$ measured by 
\cite{HewettWild10} using the \ion{Mg}{2}\,2800 emission-line, and we add the median blueshift of 
\ion{Mg}{2} from the systemic to these redshifts. We reach a final sample size of 112. 
Figure~\ref{fig:experiment} summarizes the experimental design. We refer the reader to previous 
QPQ publications for the details on the redshift centroiding algorithm, data reduction, and 
continuum normalization (QPQ6; QPQ8).

As in the previous QPQ papers, we select the quasar pairs to have redshift difference 
$>3000\,{\rm km\,s^{-1}}$, to exclude physically associated binary quasars. The cut on velocity 
difference is motivated by the typical redshift uncertainty of $\approx500\,{\rm km\,s^{-1}}$ of 
the background quasars. In QPQ8, it was required that the observed wavelengths of the 
metal ion transitions fall outside the Ly$\alpha$ forest of the background quasar. In this paper, 
we exclude a small window around the Ly$\alpha$ emission, in additional to the Ly$\alpha$ forest, 
from analysis. For stacked profile analysis, a good estimate of 
the continuum level is necessary. In QPQ8 we found that absorption associated to 
the foreground quasar occurs within $\pm2000\,{\rm km\,s^{-1}}$ around $z_{\rm fg}$. Therefore, it 
is desirable to keep a $\approx\pm3000\,{\rm km\,s^{-1}}$ window relatively free of contamination 
from Ly$\alpha$ forest. Taking into account the redshift uncertainties, 
we decide that at least one transition among \ion{C}{2}\,1334, \ion{C}{4}\,1548, and 
\ion{Mg}{2}\,2796 at $z_{\rm fg}$ must lie redward of 
$(1215.6701+20)\times(1+z_{\rm bg})\,{\rm \AA}$, for a pair to be included in the analysis.

Furthermore, we include only those spectra with average signal-to-noise ratio (S/N) exceeding 5.5 
per rest-frame \AA \ in a $\pm3000\,{\rm km\,s^{-1}}$ window centered on the observed wavelengths 
of the metal ion transitions. This criterion is a compromise between maximizing sample size versus 
maintaining good data quality on the individual sightlines. We find that S/N $>5.5$ per rest-frame 
\AA \ is necessary for properly estimating the continuum, as well as identifying mini-broad 
absorption line systems associated to the background quasar, which will significantly depress the 
flux level. We also require that the region of the spectrum that is $\pm3000\,{\rm km\,s^{-1}}$ 
around a considered metal ion transition does not overlap with strong atmospheric ${\rm O}_2$ 
bands. The ${\rm O}_2$ A- and B-band span 7595\textendash7680\,\AA \ and 6868\textendash6926\,\AA 
\ respectively.

Table~\ref{tab:sample} lists the full QPQ9 sample. In Table~\ref{tab:summary}, we first list the 
sample size, the median $z_{\rm fg}$, and the median $R_\perp$ of the quasar pairs that survive 
the above selection criteria for \ion{C}{2}\,1334, \ion{C}{4}\,1548, and \ion{Mg}{2}\,2796 
respectively. We then provide the summary for the sub-sample with $z_{\rm fg}$ measured from 
[\ion{O}{3}].

\section{Analysis}
\label{sec:analysis}

\subsection{Stacked Profiles}
\label{sec:stacks}

\begin{figure}
\includegraphics[width=3.5in]{Figures/fig_stacks_and_fits.pdf}
%\includegraphics[width=3.5in]{f02.pdf}
\caption{Mean and median absorption centered at \ion{C}{2}\,1334, \ion{C}{4}\,1548, and 
\ion{Mg}{2}\,2796 of the foreground quasars for all QPQ9 pairs. The composites are shown in thick, 
black. Overplotted on the composites are Gaussian fits, normalized to pseudo-continua far away 
from a velocity of $0\,{\rm km\,s^{-1}}$ relative to $z_{\rm fg}$, and are shown in thin, blue. 
For the doublets, a second Gaussian with a fixed mean separation and a tied standard deviation is 
included in the modeling. The absorptions frequently exhibit large velocity widths. The dashed 
lines mark the centroids, which show small positive velocity offsets from $z_{\rm fg}$. The 
$1\sigma$ modeling error of the centroid and dispersion of the \ion{C}{2} mean stack is 
$\approx35\,{\rm km\,s^{-1}}$, which is also the typical modeling errors of the other stacks. 
}
\label{fig:stacks_and_fits}
\end{figure}

We create composite spectra that average over the intrinsic scatter in quasar environments, 
continuum placement errors, and redshift errors. The individual spectra of background quasars are 
shifted to the rest-frame of the foreground quasars at the transitions of interest. Each spectrum 
has been linearly interpolated onto a fixed velocity grid centered at $z_{\rm fg}$ with bins of 
$100\,{\rm km\,s^{-1}}$. For a velocity bin of this size, it is unnecessary to smooth the data to 
a common spectral resolution. The individual spectra are then combined with a mean or a median 
statistic. Bad pixels in the individual spectra have been masked before generating the 
composites. Since each quasar pair gives an independent probe of the CGM, each pair has an equal 
weighting in the stacked profiles, i.e.\ we do not weight the spectra by the measured S/N near the 
metal ion transitions. Scatter in the stacked spectra is dominated by randomness in the CGM rather 
than scatter in the flux of individual observations. The mean statistic of the individual spectra 
yields a good estimate of the average absorption and preserves equivalent width. The median 
statistic is less sensitive to outliers, however the averaged equivalent width is smaller and the 
analysis on the velocity field becomes more uncertain. In the following we present stacked spectra 
using both the mean and the median statistic.

In Figure~\ref{fig:stacks_and_fits}, we present mean and median stacks of \ion{C}{2}\,1334, 
\ion{C}{4}\,1548, and \ion{Mg}{2}\,2796 absorption of the QPQ9 sample. We focus on the 
analysis results of the \ion{C}{2} mean stack. \ion{C}{4} and \ion{Mg}{2} are doublet transitions 
and it is more challenging to interpret their kinematics.
%To assess the average absorption of these transitions,
%we have combined the data with a novel technique.
%We have fitted the combined dataset using a cubic b-spline
%with knots at every 100\,kms.  We weighted each data
%point by the velocity width of the pixel to insure that 
%each sightline makes an equal contribution to the stack.
%These b-spline profiles are presented in the figure
%and a trivial integration gives the reported equivalent
%widths.
%[consider down-weighting systems with $<3\sigma$ EW]
%[Comnent on EW from 'standard' technique]
Two results are evident in Figure~\ref{fig:stacks_and_fits}: (i) the mean \ion{C}{2} stack 
exhibits excess absorption spanning a large velocity width; (ii) the mean absorption is likely 
skewed toward positive velocities. To model the absorption, we introduce a Gaussian profile while 
allowing a constant pseudo-continuum level to vary. We perform $\chi^2$ minimization with each 
channel given equal weight. From the best-fit to the data, we measure the centroid of the 
\ion{C}{2} stack to be $+208\,{\rm km\,s^{-1}}$ and the $1\sigma$ dispersion to be 
$312\,{\rm km\,s^{-1}}$. The centroid suggests an asymmetry that contradicts the standard 
expectation, while the dispersion suggests extreme kinematics. The median stack, on the other 
hand, shows weaker absorption, and the Gaussian model has a centroid with smaller offset and 
comparable dispersion.

We also create mean and median stack for the sub-sample with [\ion{O}{3}] redshifts and model the 
absorption with Gaussian best-fit. The \ion{C}{2} mean stack for this sub-sample has a centroid at 
$+239\,{\rm km\,s^{-1}}$, and a dispersion of $349\,{\rm km\,s^{-1}}$, consistent with the full 
sample.

To model the mean and median absorption of \ion{C}{4}\,1548 and \ion{Mg}{2}\,1796, we introduce a  
second Gaussian with separation equal to the doublet separation ($498\,{\rm km\,s^{-1}}$ and 
$769\,{\rm km\,s^{-1}}$ respectively), and tie the dispersion of the two lines in a doublet. The 
modeling results show that the velocity fields of \ion{C}{4} and \ion{Mg}{2} are consistent with 
\ion{C}{2}, i.e.\ large dispersion and centroid is skewed toward positive velocities.

The above analyses are summarized in Table~\ref{tab:summary}. The 
Gaussian models normalized to pseudo-continuum are overplotted on the data stacks in 
Figure~\ref{fig:stacks_and_fits}.

\input{Tables/tab_summary.tex}
%\input{tab_summary.tex}

\subsection{Interpretation of the asymmetric absorption}
\label{sec:significance_+ve}

Given the small sample size, one must scrutinize the statistical significance of the measured 
velocity offsets. To assess the statistical significance, we perform a boostrap analysis by 
randomly resampling from the full sample 10000 times. We introduce a Gaussian absorption profile 
to model each bootstrap realization, and quote the standard deviation in the the bootstrap 
realizations to be the scatter in the centroids of the data. The scatters 
$\approx100\,{\rm km\,s^{-1}}$ are comparable to the measured offsets, indicating large intrinsic 
variation in quasar CGM environments (see also QPQ8). In Figure~\ref{fig:histogram_cen}, we show 
the distribution of the absorption centroids from bootstrapping on the \ion{C}{2} mean stack. We 
find that 98\% of the centroids are at positive velocities. We place a generous $3\sigma$ upper 
limit to the small offset of the centroid from $z_{\rm fg}$ for the \ion{C}{2} mean absorption at 
$\delta v<+451\,{\rm km\,s^{-1}}$. Given the large intrinsic scatter, we do not attempt to explore 
whether there exists relative asymmetry among the \ion{C}{2}, \ion{C}{4}, and \ion{Mg}{2} 
absorption.

One may also ask whether the measured positive offsets come from systematic bias in redshift 
measurements due to the Baldwin effect \citep{Baldwin77}. \cite{Shen16} reported that, the 
[\ion{O}{3}] emission of $z\sim2$ quasars is more asymmetric and weaker than that in typically 
less luminous low-$z$ quasars. To test for this potential source of bias, we create another mean 
stack at \ion{C}{2}\,1334 by replacing the [\ion{O}{3}] redshifts by a \ion{Mg}{2} or H$\alpha$ 
redshift when available. We are able to replace for 8 out of the 14 systems with [\ion{O}{3}] 
redshifts in the original sample. The new stack is similiar in velocity structure and again shows 
a positive offset $\approx+240\,{\rm km\,s^{-1}}$. We thus conclude that our algorithm for 
measuring redshifts is not severely biased by the blue wing of the [\ion{O}{3}] emission-line.

Motivated by the study of \ion{Mg}{2} absorbers surrounding $z\sim1$ quasars by \cite{Johnson+15}, 
we also generate a mean-stacked spectrum for \ion{Mg}{2}\,2796 for lower redshift quasar pairs.  
We select quasars with $0.4<z_{\rm fg}<1.6$ and use the same other selection criteria as the main 
QPQ9 sample. The quasar pairs are selected from the igmspec database, with $z_{\rm fg}$ measured 
by \cite{HewettWild10}. For quasars with $z_{\rm fg}<0.84$, as redshift determination is dominated 
by [\ion{O}{3}] emission, a shift of $+48\,{\rm km\,s^{-1}}$ is applied to bring the emission-line 
redshift to the systemic. For quasars with $0.84<z_{\rm fg}<1.6$, as redshift determination is 
dominated by \ion{Mg}{2} emission, a shift of $57\,{\rm km\,s^{-1}}$ is applied. There are 231 
pairs selected, with a median $z_{\rm fg}$ of 0.90 and a median $R_\perp$ of 208\,kpc. We present 
the mean stack in Figure~\ref{fig:stack_z1}. The absorption is weaker than the $z\sim2$ main QPQ9 
sample. Gaussian absorption models fitted to the stack recover a centroid of 
$-9\pm388\,{\rm km\,s^{-1}}$ and a dispersion of $161\,{\rm km\,s^{-1}}$. The average offset 
from $0\,{\rm km\,s^{-1}}$ is much smaller than the offsets in the $z\sim2$ sample.

Since the large scatter in the centroids represents intrinsic variation rather than redshift 
errors, and the \ion{Mg}{2} stack for lower redshift suggests a different centroid, we 
consider that the asymmetry signal in the $z\sim2$ sample is likely to be real. In the Discussion 
section, we discuss two possible explanations for the asymmetry.

\begin{figure}
\includegraphics[width=2.5in]{Figures/fig_histogram_cen.pdf}
%\includegraphics[width=2.5in]{f03.pdf}
\caption{Histogram of the absorption centroids of 10000 bootstrap realizations of the data sample 
for the \ion{C}{2} mean stack. 98\% of the centroids are positively offset from $z_{\rm fg}$.}
\label{fig:histogram_cen}
\end{figure}

\begin{figure}
\includegraphics[width=2.5in]{Figures/fig_stack_z1.pdf}
%\includegraphics[width=2.5in]{f04.pdf}
\caption{Mean stack at \ion{Mg}{2}\,2796, for a lower redshift sample of quasar pairs at 
$z\sim0.9$. Line-style coding is the same as in Figure~\ref{fig:stacks_and_fits}. The centroid is 
approximately at $0\,{\rm \,km\,s^{-1}}$, and the absorption is weaker than the main QPQ9 sample. 
}
\label{fig:stack_z1}
\end{figure}

\subsection{Interpretation of the large velocity fields}
\label{sec:significance_width}

\begin{figure}
\includegraphics[width=2.5in]{Figures/fig_monte.pdf}
%\includegraphics[width=2.5in]{f05.pdf}
\caption{The same \ion{C}{2} mean stack shown in the first panel of 
Figure~\ref{fig:stacks_and_fits} is shown in thick, black. In thin, green, we overplot the 
Gaussian absorption model of the Monte Carlo simulations generated from a purely clustering 
argument. The model from clustering analysis is multiplied to the pseudo-continuum level of the 
stack of the observational data, shifted to the centroid of the stack of the data, and broadened 
by the mean redshift error in the data. The model is a good description of the data, with a 
dispersion within modeling error of the dispersion in the data. We overplot in dashed, yellow an 
absorption profile with $\sigma_v=654\,{\rm km\,s^{-1}}$, that is ruled out. Motions that produce 
this width are ruled out at $3\sigma$ confidence by bootstrap analysis on the observational data. 
}
\label{fig:monte}
\end{figure}

Under the assumption that the intrinsic dispersion and the redshift uncertainty add in 
quadrature to give the observed width, we solve for the intrinsic dispersion in the \ion{C}{2} 
mean stack. For the full QPQ9 sample, with the mean 
$\sigma_{{\rm error}(z)}^{\rm full}=182\,{\rm km\,s^{-1}}$, we recover 
$\sigma_{\rm intrinsic}^{\rm full}=253\,{\rm km\,s^{-1}}$. For the sub-sample with [\ion{O}{3}] 
redshifts, we recover $\sigma_{\rm intrinsic}^{\rm [OIII]}=342\,{\rm km\,s^{-1}}$.

\cite{Eftekharzadeh+15} measured the clustering of quasars in the range $2.2<z<2.8$ while 
\cite{RodriguezTorres+16} measured the clustering of quasars in the range $1.8<z<2.2$. They 
estimated that these quasars are hosted by dark matter halos with mass $10^{12.5}\,M_\odot$ and 
$10^{12.6}\,M_\odot$ respectively. If QPQ dark matter halos have a characteristic mass 
$10^{12.5}\,M_\odot$ and follow an NFW profile \citep{NavarroFrenkWhite97} with concentration 
parameter $c=4$, at $z\approx2$ the maximum circular velocity is $319\,{\rm km\,s^{-1}}$. 
According to \cite{Tormen97}, the maximum circular velocity is $\approx1.4$ times larger than the 
one-dimensional velocity dispersion $\sigma_{\rm 1D}$. Hence, the line-of-sight velocity 
dispersion typical of QPQ halos is $\sigma_{\rm 1D}=228\,{\rm km\,s^{-1}}$. We can investigate 
whether gravitational motions and Hubble flows are sufficient to reproduce the dispersion in the 
data, using Monte Carlo methods to simulate the absorption signals.

Since \ion{C}{2} systems arise in optically thick absorbers, we may adopt the clustering analysis 
results of QPQ6. In the absence of clustering, the expected number of absorbers per unit redshift 
interval for Lyman limit systems, super Lyman limit systems, and damped Ly$\alpha$ systems are 
respectively $\ell_{\rm IGM}^{\rm LLS}(z)\approx1.05((1+z)/(1+2.5))^{2.1}$, 
$\ell_{\rm IGM}^{\rm SLLS}(z)\approx0.44((1+z)/(1+2.5))^{2.1}$, and 
$\ell_{\rm IGM}^{\rm DLA}(z)\approx0.2((1+z)/(1+2.5))^{2.1}$. The quasar-absorber correlation 
functions for Lyman limit systems, super Lyman limit systems, and damped Ly$\alpha$ systems are 
respectively $\xi_{\rm QA}^{\rm LLS}(r)=(r/(12.5\,h^{-1}\,{\rm Mpc}))^{-1.68}$, 
$\xi_{\rm QA}^{\rm SLLS}(r)=(r/(14.0\,h^{-1}\,{\rm Mpc}))^{-1.68}$, and 
$\xi_{\rm QA}^{\rm DLA}(r)=(r/(3.9\,h^{-1}\,{\rm Mpc}))^{-1.6}$. 
For each quasar pair, we calculate the expected number of optically thick absorbers within 
$\pm3000\,{\rm km\,s^{-1}}$ at a distance $R_\perp$ from the foreground quasar and at 
$z_{\rm fg}$. Then we generate 1000 mock sightlines. The number of absorbers for each mock 
spectrum is randomly selected from a Poisson distribution with mean equal to the expected number 
calculated as above. The absorbers are randomly assigned Hubble velocities, with a probability 
distribution according to the quasar-absorber correlation functions. The absorbers are randomly 
assigned additional peculiar velocities drawn from a normal distribution with mean equal to 
$0\,{\rm km\,s^{-1}}$ and scatter equal to $\sigma_{\rm 1D}$. For each absorber, we assume a rest 
equivalent width for \ion{C}{2} $W_{\rm CII}$ and a Gaussian absorption profile. We repeat the 
above procedure for all 32 quasar pairs, and create a mean stack of the 32000 mock spectra 
generated. We fit a Gaussian absorption profile multiplied to a constant continuum level to model 
the stack of mock spectra. We adjust the $W_{\rm CII}$ adopted for the absorbers until the 
amplitude of the best-fit Gaussian of the stack of mock spectra matches the 
amplitude of the stack of the observational data. We find that $W_{\rm CII}=0.7\,{\rm \AA}$ well 
reproduces the amplitude, and the dispersion of the Gaussian absorption model is 
insensitive to the assumed $W_{\rm CII}$ or line profile for one absorber.

In Figure~\ref{fig:monte}, we show a comparison of the observational data stack and the Gaussian 
absorption model of the Monte Carlo simulations. In the figure, the Gaussian absorption model is 
broadened by the mean redshift error by adding it in quadrature to the dispersion in the model. 
The resulting stack of mock spectra has a $1\sigma$ dispersion of $263\,{\rm km\,s^{-1}}$. This is 
within $1\sigma$ modeling error of the intrinsic dispersion in the \ion{C}{2} mean stack for the 
full sample, and about 2 times the modeling error away from the intrinsic dispersion in the stack 
of the sub-sample with [\ion{O}{3}] redshifts. We also test for the 
sensitivity of this measured dispersion to the correlation functions adopted. The QPQ6 clustering 
analysis is performed on only the strongest absorber near $z_{\rm fg}$, and a low $R_\perp$ 
sightline may in fact intercept more than one optically thick absorber. We double the number of 
absorbers for each mock sightline, and find the measured dispersion only increases by several 
${\rm km\,s^{-1}}$. Thus, the hypothesis that the observed velocity width is only produced by a 
combination of gravitational motions and Hubble flows cannot be ruled out. Extra dynamical 
processes are not necessary to explain the large velocity fields. From the bootstrap analysis of 
the observed dispersion, a width of $\sigma_v>654\,{\rm km\,s^{-1}}$ is ruled out at 3$\sigma$ 
confidence (overplotted in Figure~\ref{fig:monte}). In other words, motions additional to 
gravitational and Hubble flows that will produce intrinsic dispersion 
$\sigma_v>628\,{\rm km\,s^{-1}}$ are ruled out.

On the contrary, $\sigma_{\rm 1D}<175\,{\rm km\,s^{-1}}$ together with Hubble velocities will 
result in a velocity width that is more than three times the modeling error away from the observed 
intrinsic width. This implies that if the characteristic dark matter halo mass 
$<10^{12.2}\,M_\odot$, additional dynamical processes are required to explain the observed width.

\subsection{Line-of-sight absorption}
\label{sec:los_stacks}

\begin{figure}
\includegraphics[width=3.4in]{Figures/fig_stacks_fg.pdf}
%\includegraphics[width=3.4in]{f06.pdf}
\caption{Mean stacks of the foreground quasar spectra at \ion{C}{2}\,1334, \ion{C}{4}\,1548, and 
\ion{Mg}{2}\,2796 for the QPQ9 sample. For \ion{C}{2} and \ion{Mg}{2}, the mean absorption is 
weaker than that in the background stacks, and there is no evidence for an excess at $z_{\rm fg}$. 
The stack for \ion{C}{4}, which includes line-of-sight absorbers at all distances, shows a large, 
blueshifted mean velocity field.
}
\label{fig:stacks_fg}
\end{figure}

In the previous QPQ papers, we argued that optically thick absorbers in the vicinity of quasars 
are distributed anisotropically. We now have the means to show this anisotropic clustering 
explicitly. Given that the techniques for stacking spectra are established, it is straightforward 
to apply the same techniques to stack the foreground quasar spectra. In 
Figure~\ref{fig:stacks_fg}, we present mean stacks of foreground quasar spectra for 
the QPQ9 sample. We require that the spectra survive a S/N cut of 5.5 per rest-frame \AA\ at 
\ion{C}{2}\,1334, \ion{C}{4}\,1548, or \ion{Mg}{2}\,2796 at $z_{\rm fg}$. In constrast to the 
large equivalent widths exhibited in the stacks of background spectra, \ion{C}{2} and \ion{Mg}{2} 
mean absorption along the line-of-sight to the foreground quasars is weaker, and an excess at 
$z_{\rm fg}$ is absent. This supports a scenario where the ionizing radiation of the foreground 
quasars are anisotropic and/or intermittent. For \ion{C}{4}\,1548, this stack of all line-of-sight 
absorbers, which include absorbers intrinsic to and far away from the quasar, shows a large, 
blueshifted velocity field. This excess \ion{C}{4} absorption has been well studied as narrow 
associated absorption line systems \citep[e.g., ][]{Wild+08}.

\section{Discussion}
\label{sec:discussion}

\begin{figure*}[!th]
\begin{minipage}[!bp]{0.28\textwidth}
\includegraphics[width=\textwidth]{Figures/fig_monopolar.pdf}
%\includegrpahics[width=\textwidth]{f07.pdf}
\caption{A cartoon showing a unipolar quasar. The gas observed in low- to intermediate-ion 
absorption preferentially lies behind the quasar, and is shadowed from the ionizing radiation. 
}
\label{fig:monopolar}
\end{minipage}%
\hspace{0.3in}
\begin{minipage}[!bp]{0.67\textwidth}
\includegraphics[width=\textwidth]{Figures/fig_lighttravel.pdf}
%\includegraphics[width=\textwidth]{f08.pdf}
\caption{A cartoon showing the finite lifetime of quasar episodes as an explanation to the 
asymmetric absorption. The setup on the left shows that the foreground quasar has not been shining 
long enough for its ionizing radiation to reach the gas behind it, when the light from the 
background quasar reaches. The setup on the right shows the scenario after an amount of time 
comparable to the light travelling time across CGM scale. Gas in front of the foreground quasar 
has been ionized, by the time the light from the background quasar reaches. 
}
\label{fig:lighttravel}
\end{minipage}
\end{figure*}

In the following, we explore two possible explanations for the putative non-dynamical process that 
provides the asymmetry.

One possibility is an asymmetric radiation field that preferentially ionizes the gas moving toward 
the observer, where the quasar is known to shine. Alternatively, the asymmetric radiation field 
may preferentially ionizes the gas at smaller Hubble velocity than the quasar. In 
Figure~\ref{fig:monopolar}, we show a cartoon of a quasar that is blocked in the direction 
pointing away from the observer. The gas observed in absorption preferentially lies behind the 
quasar. The asymmetric absorption arises from a transverse proxmity effect. 
\cite{Roos+15} and \cite{GaborBournaud14} performed simulations of a high-redshift disk galaxy 
inlcuding thermal AGN feedback and calculated radiative transfer in post-processing. They found 
the ionization radiation is typically asymmetric, due to either a dense clump that lies on one 
edge of the black hole or the black hole's location being slightly above the disk.

Another possible explanation arises from the finite lifetime of quasar episodes. 
Figure~\ref{fig:lighttravel} presents a cartoon for a light travel time argument. Light from the 
background quasar will travel smaller distance to reach the gas behind the foreground quasar 
than the gas in front of it. The light from the background quasar may arrive at the gas behind 
the foreground quasar before the ionizing radiation from the foreground quasar arrives. In 
QPQ7 and QPQ8, we reported that the enhancement in metal ion absorption relative to IGM average 
extends to at least $\approx200$\,kpc. This distance corresponds to a light travelling time of 
$\sim10^6$\,Myr, which lies within existing constraints of quasar lifetime from observations 
\citep[e.g.][]{Martini04} and simulations \citep[e.g.][]{Hopkins+05}.

As a test of reasonableness, we calculate that, at $R_\perp\approx50\textendash150$\,kpc, radial 
speeds relative to the foreground quasar of $500\textendash700\,{\rm km\,s^{-1}}$ and quasar 
opening angles of $120^\circ\textendash140^\circ$ may reproduce the observed centroid and 
intrinsic dispersion of the \ion{C}{2} mean absorption. The first explanation above to asymmetric 
absorption requires that the quasars emit their ionizing radiation anisotropically, while the 
second explanation requires the quasars emit their ionizing radiation intermittently. Both 
explanations will require the gas to be either outflowing or in Hubble flow. Were the gas 
inflowing instead, the velocity centroid would be negative.

Motivated by the asymmetry found in metal ion absorption in the CGM using precise $z_{\rm fg}$ 
measurements, and the asymmetry found by \cite{KirkmanTytler08} in \ion{H}{1} on larger 
scales, we are assembling a sample of quasar pairs with precise $z_{\rm fg}$ measurements to study 
this asymmetry in \ion{H}{1} (J. F. Hennawi et al. 2017, in preparation). In conclusion, we 
observe large and positively skewed velocity fields in absorption, of metal ions 
in the CGM of $z\sim2$ massive galaxies hosting quasars. We argue that the velocity fields can be 
explained if the detected gas is in gravitational motion, Hubble flow, or outflow, and the quasars 
either shine preferentially toward the observer or shine intermittently.

\acknowledgements

JXP and MWL acknowledge support from the National Science Foundation (NSF) grants AST-1412981. The 
authors gratefully acknowledge the support which enabled these observations at the Keck, Gemini, 
Large Binocular Telescope, Very Large Telescope, Las Campanas, and Palomar Observatories. MWL 
thanks T.-K. Chan for a discussion on CGM simulations and Chi Po Choi for a discussion on 
statistics.

\bibliographystyle{/Users/lwymarie/Documents/Bibli/apj}
\bibliography{/Users/lwymarie/Documents/Bibli/allrefs}

\end{document}