### 2017-07-20 QPQ9/Submit/response.txt
Responded to referee

Editor Walter,

We thank very much the referee for their careful reading of our manuscript, and their comments and criticism.
Following their recommendations, we have revised our manuscript in a number of ways detailed below. We are now
resubmitting to the Astrophysical Journal and ask that you consider it further for publication.
Summary of referee report:

>After carefully reviewing the manuscript, I found multiple significant issues with the study, and consequently it does
>not currently meet the standards for publication. These issues must be addressed before the paper can be accepted, and
>I request the opportunity to review the paper again after the authors have made revisions with the hope that it can be
>accepted for publication. The main issues are summarized below and more detailed discussion follows.
>
>1) The conclusions of the paper rely heavily on accurate quasar redshift measurements because redshift errors will
>influence both the centroid and dispersion measured in the stacked background absorption spectra. Indeed, the measured
>centroids and velocity dispersions in the stacked absorption spectra are comparable with expected quasar redshift
>biases and scatter. In an attempt to address this issue, the authors adopt redshift bias measurements from the
>literature, but the bias measurements used are not appropriate for the redshift measurement techniques that are
>described in the previous QPQ papers (details to follow). Significant additional work is required to demonstrate that
>the offset in the observed absorption centroid is real and not due to redshift measurement errors.

We thank the referee for pointing out the importance of accurate redshift measurements. We have addressed the
suggested additional work in the details to follow.

>2) The authors have not demonstrated that the measured 1-sigma dispersion measured in a mean stack is is a
>representative measure of the velocity spread observed between absorbing gas and the quasar redshift. This may not be
>the case if the absorbing components exhibit a non-Gaussian distribution about the quasar systemic redshift as has
>been seen in other studies (e.g. Lau et al. 2016 and Johnson et al. 2015). In particular, broad wings on the
>distribution will be suppressed in a stack even if they contain a significant amount of the power because the
>absorbing components can be narrow. This is a weakness of analysis based on stacks that could easily be remedied
>because previous QPQ papers published absorber redshifts and even component-by-component velocity structure when
>available. A figure showing this distribution is needed in this publication.

The distribution of absorbing components in velocity is non-Gaussian for each quasar-host associated system. However,
we have 34 absorption systems for CII analysis, and an even larger number of absorption systems for CIV and MgII
analysis. The number of individual components can only be larger. The central limit theorem applies and the
distribution of all absorbing components should tend toward Gaussian distribution. A component-by-component analysis
requires echellette or echelle resolution, which is a criterion for the QPQ8 sample. For this current work, we stack
spectra of all resolutions. Text has been added to explain the above and to refer the readers to specific figures in
QPQ8.

>3) Insufficient detail on the doublet fitting is provided, and uncertainty due to degeneracy between doublet ratio and
>velocity centroid are not addressed.

We allowed the doublet ratio to vary from 2:1 to 1:1. The centroids of the doublets are tied at a fixed separation.
More text has been added.

>4) The conclusions of the paper are not self-consistent. In particular, the authors find that the velocity dispersion
>in the mean stack does not require addition motion (beyond gravitational and cosmological), but the explanation for
>the centroid offset requires that the gas is, on average, flowing away from the quasar host galaxy. This discrepancy
>needs to be addressed.

We acknowledge that we did not present the best argument. The centroid offset does not require that the gas is
flowing away from the galaxy. In our Monte Carlo simulations, we are able to reproduce the centroid offset by assuming
that the gas is all ionized within a certain opening angle. For an opening angle >= 120 degrees, The stack of mock
spectra has amplitude, dispersion, and centroid offset all within 2 times the modeling error. Text has been added to
explain the above. We do not present a new figure to show this stack, since our primary goal of performing Monte Carlo
simulations is to reproduce the dispersion. We have removed the outward arrows in the cartoon figure and the estimates
in radial velocities under an outflow scenario. While outflows cannot be ruled out, we prefer to focus on the argument
that outflows are not required.

>5) Several papers in the literature that are cited are mis-represented.

We have studied the papers more carefully, and have modified the text to represent the literature more correctly.
Details to follow below.

>Additional details on each of the points as well as more minor issues are given below, After this, my referee report
>concludes with a discussion of more minor points.
>
>Major point 1: quasar systemic redshift measurements
>
>The authors compile a foreground quasar sample by combining quasars with redshifts measured in their own spectra with
>quasars from the SDSS with redshift measurements from the literature. Very little information is provided on how these
>redshifts are measured and the reader is referred to previous QPQ papers. Given the importance of accurate redshifts
>for this study, this lack of discussion is inappropriate. The authors provide significant detail in how the redshifts
>are measured and demonstrate the accuracy of their redshifs.
>
>After returning to the QPQ papers where redshift measurements are discussed, it appears that the QPQ redshifts are
>based on *centroids* from MgII, [O III], Halpha emission lines. The authors then apply mean bias corrections for these
>redshifts based on the line-dependent bias measured by Shen et al. 2016a. However, Shen et al. 2016a measured the bias
>for redshifts based on line *peak*, not centroid. This may seem subtle but is significant because of the strong
>blue-shifted wings often observed in quasar emission-lines. For example, Shen et al. 2016b fit z~2 quasar emission
>lines with a core component and broad wing component. The broad wings have a mean velocity offset of -600 km/s from
>the systemic redshift and contain 40% of the [O III] flux on average. Consequently, the [O III] centroid will be
>biased from the systemic redshift measured from the peak by ~-200 km/s which is sufficient to explain the redshift
>observed in the stacked absorption profile. Note that this is far larger than the 50 km/s correction that the authors
>applied which was based on peaks not centroids. Redshifts from MgII and Balmer emission line centroids are likely even
>more biased because these lines are broader and more asymmetric. Consequently, the authors must re-measure the quasar
>systemic redshifts using techniques that are self-consistent with the redshift corrections that are applied.
>
>In addition, the authors need to make and display stacked absorption profiles in each ion for quasars with different
>redshift measurement techniques. Some effort along these lines is shown in Table 2, but it is not nearly sufficient to
>demonstrate that the measured absorption centroid is not due to redshift measurement issues.
>
>These steps are a minimum requirement to convincingly demonstrate that the observed redshift in the mean absorption
>profile relative to the quasar systemic is real.

We have added text to explain how redshifts are measured. We acknowledge the confusion in calling the line-centering
algorithm a centroiding method. Our algorithm calculates the mode of a line given by (3*median - 2*mean), applied to
the upper 60% of the emission line. Shen+16 states that difference between the peak and the centroid is not
significant, except for the broad line Hbeta, which we did not use in redshift measurements. We have further obtained
individual measurements of centroids and peaks in Shen+16 thru private communication. There is essentially no
difference between using the centroid versus using the peak for [OIII], and there is about 50 km/s difference for
MgII. We may expect the difference between a mode and a peak is even smaller. Text have been added.

We have explained in the text that, our results cannot be severely biased by [OIII] redshifts. We have tried
replacing the [OIII] redshifts by MgII or Halpha redshifts, and found a similar centroid and dispersion in the stack.
MgII and Halpha are broader, but it has not been reported in any literature that they are more asymmetric.

>Major point 2:
>
>In Lau et al. 2016, the distribution of absorbing components about the quasar systemic redshift is non-Guassian and
>exhibits broad wings that contain significant portion of the optical depth at large velocities. In a mean stack, these
>wings can be suppressed so that the dispersion of the stacked absorption profile does not represent the actual
>velocity spread. For example, if absorbers are distributed in velocity around quasar hosts with a core component
>within |v|<200 km/s and a wing component within |v|=200-1000 km/s, then the wing component will be suppressed by a
>factor of ~8 relative to the core because narrow absorbers are being spread over a larger velocity range. The authors
>need to demonstrate that this is not an issue before adopting the 1-sigma dispersion in the stack as a measure of the
>kinematic spread of the absorbing gas. This is a general weakness of stacked measurements and raises the question of
>why the study does not show a histogram of the velocity difference sbetween the quasar systemic and observed
>absorption centroids in individual cases where the S/N is sufficient. Previous QPQ papers reported absorber redshifts
>(Prochaska et al. 2014) and individual component measurements (Lau et al. 2016). It should therefore be
>straight-forward to measure the actual velocity distribution of absorption systems around quasars directly rather than
>relying on stacked measurements. The authors should include such a histogram in the revised paper.

As explained above, we have sufficient number of absorbing components that they should tend toward a Gaussian
distribution in the stack. We are able to model the stack by a Gaussian multiplied to a pseudocontinuum. Our visual
inspection does not hint at a broader component in the stack. QPQ7 reported the redshift of the strongest Lya
absorption within the search window. The QPQ7 sample has spectra of all resolutions. It requires echellette or
echelle resolution to identify the metal-associated absorbers. In QPQ8, a histogram of the optical depth weighted
centroid of the absorption systems is shown. In the Appendix of QPQ8, the absorption profiles of the individual
components are shown. We have counted 6 of them are at negative velocities, while 15 of them are at positive
velocities. text have been added to reflect these numbers.

>Major point 3:
>
>The authors do not provide enough detail regarding the fit to the mean stack. In particular, for MgII and CIV two
>gaussian with fixed doublet separation and width are fit to the stack. However, no details on the doublet ratio range
>allowed are given. This is particularly important for CIV where the doublets are not well resolved in the stack.
>Consequently, the velocity centroid and doublet ratio will be degenerate and uncertainty due to this degeneracy must
>be addressed.

We thank the referee for pointing out the insufficient detail in the text. The doublet ratio is allowed to vary
between 2:1 to 1:1.

>Major point 4:
>The primary results of the paper are that (1) while the dispersion observed in the stack is large, galactic superwinds
>may not be required to explain the width and (2) the absorption is kinematically offset from the quasar systemic
>redshifts by ~+200 km/s. The authors interpret this offset as due to effects from partial obscuration of the quasar or
>finite quasar lifetimes and light-travel time under the assumption that gas is, on average, flowing away from the
>quasar host. In QPQ VII the CII covering fractions are observed to decrease rapidly from ~80% at projected distances
>of <200 kpc to 20% at 200-500 kpc and <10% at 500-1000 kpc. This rapid fall-off with projected distance suggests that
>the majority of the absorption arises in the quasar host halo. Consequently, motions due to Hubble flow are not likely
>to be significant. Conclusion (2) therefore requires significant outflows in the halo, inconsistent with conclusion
>(1). This discrepancy needs to be addressed in the discussion.

In QPQ8, we have estimated the random incidence of optically thick and CIV absorbers. Text has been added to reflect
these numbers. As our Monte Carlo simulations demonstrate, outflows are not required to reproduce the observed
amplitude and dispersion of the mean absorption.

>Major point 5:
>
>Several papers in the literature that are cited are mis-represented in the introduction. In particular:
>
>Johnson et al. 2015 is cited as concluding that the large velocity of absorbing gas observed around quasars is due to
>outflows. This is not the case. The paper considers several possible explanations including large-scale structure,
>galaxy interactions, and outflows. The conclusion of the paper is that none of these possibilities can satisfactorily
>explain the observed properties of cool gas around quasars.
>
>Gauthier et al. 2013, Muzahid 2015, and Zahedy et al. 2016 exhibit velocity spreads of ~600, 300-400, and 600 km/s
>respectively, which is comparable with those observed in the QPQ survey, contrary to the claim in the paper. This
>inaccurate statement needs to be clarified and corrected.
>
>These misrepresentations combined with the fact that the paper applies redshift bias corrections from the literature
>that are not appropriate for their redshift measurement techniques indicates that the authors need to carefully review
>the paper and ensure that the citations to the literature are correctly represented and applied.

We thank the referee for clarifying the explanation to Johnson+15's results. We have updated the text to reflect that
outflows is only one possibility. Gauthier+13 is a single sightline, and the velocity width Dv90 of the MgII
absorber is less than the average of QPQ8 CII absorber. Muzahid+15 found an absorber with Dv90 less than 555 km/s in
OVI and NV, and still less for other ions. Zahedy+16's sample has average Dv90 elss than that of QPQ8. More text has
been added.

>Minor issues:
>
>Introduction Section:
>
>1) When discussing inflows, the authors should cite Rubin et al. 2012
>2) The definition of h in this paper is not in accordance with the common convention and should be replaced with h_70
>to prevent the casual reader from being confused.
>3) Rudie et al. 2017 should be included in the references for systems with large velocity spread. This paper is
>highly relevant to the discussions in this manuscript.

We have cited Rubin+12 and Rudie+17 in our revised manuscript. We intend that h = H_0 / (100 km s^-1 Mpc^-1), in
accordance with the common convention, that is also widely used in cosmology papers such as Kravtsov+17. We have
state this scaling explicitly in our revised manuscript. We thank the referee for above suggestions.

>Experiment Section:
>
>1) Given the anti-correlation between covering fraction and projected distance, the authors may be able to increase
>the S/N in the average absorption profiles by only considering quasars with projected distances of <100 and <200 kpc
>from the background sightline.

We performed the experiment of separating the sightlines into two bins of projected distance. We found the stack of
sightlines below the median projected distance looks similar to the full sample. We did not find a significant boost
in S/N in the average absorption profiles.

>Analysis Section:
>1) The authors claim that it is not necessary to smooth the spectra before linearly interplating onto a 100 km/s
>pixel grid. This may be true for the low-resolution spectra but is untrue for the moderate and high resolution
>spectra used where linear interpolation onto a coarser pixel scale will have unpredictable results. The authors
>should test that this dues not actually change any results.

For each sightline, we overplotted the original absorption profile and that interpolated onto the common velocity
grid, and visually the overplots. We have confirmed to ourselves that the equivalent widths and the shapes of the
absorption profiles are preserved.

>2) The mean redshifted centroid in the CII profile is observed in the mean stack but not the median at a level that
>is in tension. The study does not address this discrepancy and does not provide sufficient reasoning for preferring
>the mean over the median when attempting to measure a centroid.

There is large intrinsic scatter in the CGM environments. In our revised manuscript, we included one new quasar pair
from our latest QPQ Survey database, making the total sample size for CII analysis 33. With this sample, the median
CII statistic has a centroid +120+/-195 km/s. We prefer the mean over the median, as the median statistic is not
sensitive to absorption signals and does not preserve equivalent width. The centroid measured from the median stack
is thus subject to larger uncertainty.

>3) It is quite odd that the quoted error in the centroid for the z=1 coadd is 388 km/s, larger than the 161 km/s
>dispersion. What is going on here? In addition, why would a shift be observed for z=2 quasars but not z=1 ones? The
>authors seem to take this as evidence that the redshift in the absorption centroid is real. I came to the opposite
>conclusions since accurate redshifts for low-z quasars are easier to come by.

The quoted uncertainty in the centroid is the standard deviation in the 10000 bootstrap realizations. The average
absorption of this stack is weak, and the measured dispersion too has a large intrinsic scatter. A shift is not
definitively observed in the 0.4 < z < 1.6 stack. It may be due to weaker average absorption, or it may be due to
lower average quasar luminosity should our interpretations be correct. The 0.4 < z < 1.6 quasars are largely SDSS
spectra, and we adopt [OIII] or MgII emission-line redshifts taken from Hewett&Wild10. The z ~ 2 quasars of our sample
have MgII, [OIII], or Halpha emission-line redshifts. We do not consider that the redshifts of the 0.4 < z < 1.6
quasars are more accurate, since the more precise [OIII] redshifts are only accessible in SDSS spectral coverage for
z < 0.84.

>4) Section 3.4 contributes little to the paper and is redundant with existing work in the literature. The authors may
>wish to remove it to avoid distracting readers.

We have moved this section to the Appendix in our revised manuscript.

>5) Figure 6 may be difficult for color-blind readers to interpret.

We consulted websites for generating color-blindness-friendly figures, and found that our navy color was challenging
for readers with total color-blindness. We have changed that to a brighter shade of blue in our revised manuscript.

Discussion Section:

>1) The paper cites Roos et al. 2015 and Gabor & Bournaud et al. 2014 when discussing quasar emission anisotropy. The
>should add references to observational papers that demonstrate this though polarization measurements and/or
>observations of ionization cones at lower redshifts.

At low z, both symmetric and asymmetric ionization cones around quasars and AGNs are observed. In the extended
emission-line region of 4C37.43, most of the [OIII] emission is blueshifted (Fu&Stockton07), but there are counter
examples in less extended sources in FU&Stockton09. Recently there has been Fabry-Perot velocity fields for less
extended narrow-line regions in more nearby sources (Keel+15,17), which show mostly symmetric velocity fields and gas
distributions.

>2) In the cartoons presented in Figure 8-9, my interpretation is that under this scenario, low ions should be
>redshifted and high ions such as CIV blue-shifted. This is not observed based on the measurements of CIV in Table 2.
>Please clarify and address this apparent contradiction between the observed data and model interpretation.

CIV may be regarded as an intermediate ion, and gas that receives an ionizing radiation from the quasar may be even
more highly ionized. We have access to a Cloudy grid that uses an extragalactic UV background spectral shape as input.
We have verified to ourselves the following. For the typical Lyman limit luminosity of our sample
10^31.5 erg s^-1 Hz^-1, for the average total column density of the QPQ8 sample 10^20.5 cm^-2 and the average [M/H] =
-0.6, at a distance = 150 kpc and density = 0.1 cm^-3, Cloudy predicts that CIV absorption is only marginally
detectable. Moreover, given the large intrinsic scatter, we do not explore whether there exists relative asymmetry
among CII, CIV, and MgII.

>3) Finally, if the authors are aware of any work simulation work demonstrating that gas flows at high-z on these
>scales are primarily outward, they should be discussed. For example, Shen et al. 2012, find significant column
>densities in inflowing gas, particularly at lower ionization states. Of course, this is a lower mass halo, but
>similar insights may be available from simulations of more massive systems.

Faucher-Giguere+16 represents the only simulation so far that is able to reproduce the substantial amount of cool gas
in quasar-mass haloes. Their group has not compared the fraction of gas in inflows and outflows. Furthermore, we
consulted Jared Gabor and we found no more simulation papers that predicted unipolar emission other than the two we
cited.