### 2017-09-28 QPQ9/Submit/qpq9_submit.tex
resubmission, before bolding

%\documentclass[12pt,preprint]{aastex}
\documentclass[iop]{emulateapj}
\usepackage{natbib,amsmath,graphicx,longtable,amssymb,latexsym,epsf}
\begin{document}

\title{Quasars Probing Quasars IX. The Kinematics of the Circumgalactic Medium Surrounding 
$z\sim2$ Quasars}

\author{Marie Wingyee Lau\altaffilmark{1,2}, J. Xavier Prochaska\altaffilmark{2}, 
Joseph F. Hennawi\altaffilmark{3}
}
\altaffiltext{1}{Email: lwymarie@ucolick.org}
\altaffiltext{2}{Department of Astronomy and Astrophysics, UCO/Lick Observatory, University of 
California, 1156 High Street, Santa Cruz, CA 95064}
\altaffiltext{3}{Department of Physics, Broida Hall, University of California, Santa Barbara, 
CA 93106-9530}

\begin{abstract}
We examine the kinematics of the gas in the environments of galaxies hosting quasars at $z\sim2$. 
We employ 148 projected quasar pairs to study the circumgalactic gas of the foreground quasars in 
absorption. The sample selects foreground quasars with precise redshift measurements, using 
emission-lines with precision $\lesssim300\,{\rm km\,s^{-1}}$ and average offsets from the 
systemic redshift $\lesssim|100\,{\rm km\,s^{-1}}|$. We stack the background quasar spectra at the 
foreground quasar's systemic redshift to study the mean absorption in \ion{C}{2}, \ion{C}{4}, 
and \ion{Mg}{2}. We find that the mean absorptions exhibit large velocity widths 
$\sigma_v>300\,{\rm km\,s^{-1}}$. The observed widths are consistent with gas in gravitational 
motion and Hubble flow, and galactic-scale outflows are not required to explain the large 
widths. Furthermore, we find that the mean absorptions are asymmetric about the systemic redshift. 
The mean absorption centroids exhibit small redshift relative to the 
systemic $\delta v\approx+200\,{\rm km\,s^{-1}}$, with large intrinsic scatter in the centroid 
velocities of the individual absorption systems. We show that the observed offsets may be produced 
if (i) the ionizing radiation from the foreground quasars is anisotropic or intermittent; (ii) the 
gas is not flowing into the galaxy.  
\end{abstract}

\keywords{glaxies: clusters: intracluster medium -- galaxies: formation -- galaxies: halos -- 
intergalactic medium -- quasars: absorption lines -- quasars: general}

\section{Introduction}
\label{sec:introduction}

Galaxy formation and evolution are driven by the flows of gas into and out of their interstellar 
medium. Current theories demand that star-forming galaxies maintain these flows. Gas accretes, 
cools, and adds to the fuel supply, while star formation feedback heats gas, blows it out of 
galaxies, and regulates star formation \citep[for a review see][]{SomervilleDave15}.

Direct observations of galactic flows are difficult to acquire. Detecting the presence of 
the gas is itself challenging. Either the gas mass is too small, or the gas density is too low for 
the detection of line-emission, e.g.\ 21\,cm, Ly$\alpha$, or H$\alpha$ from \ion{H}{1}. Resolving 
the kinematics and establishing the mass flux pose an even greater challenge. These challenges are 
accentuated for distant, young galaxies, where flows of gas are predicted to prevail 
\citep{Keres+09,Fumagalli+11}. Therefore, with rare exceptions, 
\citep[e.g.][]{Cantalupo+14,Hennawi+15}, the community has relied on absorption-line spectroscopy 
to detect and characterize the gas surrounding galaxies 
\citep[e.g.][]{BergeronBoisse91,Steidel+10,Prochaska+11,Tumlinson+13}. 
In a previous paper of the Qussars Probing Quasars series \citep[][, hereafter QPQ8]{QPQ8}, we 
measured the velocity field for \ion{C}{2}\,1334 and \ion{C}{4}\,1548, finding that the 
circumgalactic medium frequently exhibits large velocity widths that are offset from the systemic 
redshift. The offsets, $\delta v$, are often positive, with the sign convention that positive 
velocities indicate a redshift from the systemic. From a sample of 7 \ion{C}{2} systems and 10 
\ion{C}{4} systems, we measured the velocity interval that encompasses 90\% of the total optical 
depth, $\Delta v_{90}$, and the $1\sigma$ dispersion relative to the profile centroid, $\sigma_v$. 
The median $\Delta v_{90}$ is $555\,{\rm km\,s^{-1}}$ for \ion{C}{2}\,1334 and is 
$342\,{\rm km\,s^{-1}}$ for \ion{C}{4}\,1548. The median $\sigma_v$ is $249\,{\rm km\,s^{-1}}$ for 
\ion{C}{2}\,1334 and is $495\,{\rm km\,s^{-1}}$ for \ion{C}{4}\,1548. These velocity fields exceed 
all previous measurements from galaxies and/or absorption systems at any epoch.

With absorption-line spectroscopy of background sightlines, other researchers also have had 
success in characterizing the flows of gas around galaxies. \cite{Rakic+12} found a net 
large-scale inflow around star-forming galaxies, or a Kaiser effect for gas on 
1\textendash2\,Mpc scales. \cite{Ho+17} found gas spiraling inward near the disk plane of 
star-forming galaxies on $<100$\,kpc scales. \cite{Johnson+15} studied the CGM surrounding 
$z\sim1$ quasars. They found large peculiar motions in the gas exceeding the expected virial 
velocity, with quasar-driven outflows being one possible explanation. However, we consider that 
their velocity spreads can be better quantified. Other 
existing studies that found large velocity spreads are limited to single sightlines with most of 
the gas within several hundred ${\rm km\,s^{-1}}$ \citep[e.g., ][]{Tripp+11,RudieNewmanMurphy17}, 
or with gas tracing 
a higher ionization state than the QPQ absorption systems \citep[e.g., ][]{Churchill+12}, or where 
the average velocity spread is smaller than that measured in QPQ8. In \cite{Gauthier13}, where a
single sightline is reported, the $\Delta v_{90}$ of the \ion{Mg}{2} absorption is less than the 
average of the QPQ8 \ion{C}{2} absorption. In \cite{Muzahid+15}, an absorber is found with 
$\Delta v_{90}$ smaller than $555\,{\rm km\,s^{-1}}$ in \ion{O}{6} and \ion{N}{5}, and still 
smaller for other ions. The \cite{Zahedy+16} sample likewise has average $\Delta v_{90}$ smaller 
than that of QPQ8.

A significant limitation of absorption-line analysis of transverse sightlines, especially 
regarding galactic-scale flows, is the inherent symmetry of the experiment. One generally lacks 
any constraint on the distance of the gas along the sightline. Positive or negative velocities 
with respect to the galaxy may be interpreted as gas flowing either toward or away from the 
system. ``Down-the-barrel'' observations break this symmetry, and have generally provided evidence 
for flows away from galaxies \citep{Rupke+05,Martin05,Weiner+09,Rubin+14}. However, these data are 
frequently at low spectral resolution which limits one's sensitivity to inflowing gas.

In this paper (hereafter QPQ9), we examine the flows of gas in the environments of massive 
galaxies hosting quasars. Our approach leverages a large dataset of quasar pairs 
\citep[][hereafter QPQ1]{QPQ1} to use the standard techniques of absorption-line spectroscopy 
with background quasars. These quasar pairs 
have angular separations that correspond to less than 300\,kpc projected separation at the 
foreground quasar's redshift. Our previous publications from these quasar pairs have established 
that these galaxies are surrounded by a massive, cool, and enriched CGM 
\citep[QPQ5, QPQ6, QPQ7:][]{QPQ5,QPQ6,QPQ7}. We have collected a sample of 102 background 
spectra that are paired with foreground quasars with precisely measured redshifts. Among the
sightlines in the QPQ9 sample, 13 have spectral resolution $R>5000$ from echellette or echelle
observations, and have been analyzed separately in QPQ8 (see their Figure 8, 9, and 10, and their 
Appendix). In QPQ8, where the individual metal-bearing absorption components are resolved, 15 out 
of the 21 components are at positive velocities relative to the systemic redshift. For this
current work, we stack spectra of all resolutions instead of performing a component-by-component
analysis. Our primary scientific interests are twofold: (i) search for signatures of galactic-scale
outflows from the central galaxy, presumably driven by recent star formation and/or active galactic
nuclei feedback; (ii) characterize the dynamics of the gas around these massive systems. We further
describe an aspect of this experiment that offers a unique opportunity to study galactic-scale
flows: we argue that, the anisotropic or intermittent radiation from the foreground quasars may
break the symmetry in the velocity field of circumgalactic absorbers. If the ionizing radiation
field is asymmetric, the absorbers may also distribute asymmetrically. Alternatively, finite quasar
lifetime will result in different radiation fields impinged on the gas closer to versus further
away from the background quasar, due to different light travel times. \cite{KirkmanTytler08}
reported an asymmetry in \ion{H}{1} absorption on scales larger than the CGM, and gave similar
arguments on anisotropy or intermittence.

%The formation and evolution of stars are dictated
%by the flow of gas and metals/dust.  [This transport
%varies directly with the stellar stage
%and a lesser extent stellar mass.]
%The accretion onto a protostellar  core eventually
%initiates nuclear fusion.  The main sequence is characterized
%by gentler stellar winds, with the most massive
%stars shedding a cumulative mass of XX\%.
%Stellar deaths, meanwhile, are frequently accompanied by terrific
%mass loss, occasionally punctuated by a highly energetic,
%supernova phase.
%By analogy, galaxy formation and evolution are driven by the
%flows into/out of their interstellar medium.
%In contrast to most stars, however, current theories
%predict (even demand) that the star-forming galaxies
%maintin these flows. [add]
%Akin to stars, direct observations of galactic flows are
%difficult to acquire. Even detecting the gas is challenging:
%the gas mass is either too small or the gas density too low
%for the [detection] of line-emission (e.g. 21cm, lya,
%or H$\alpha$ from Hydrogen).  Resolving the kinematics
%and establish the mass-flux represent a far greater challenge.
%These challenges are accentuated for distant, young galaxies
%where flows are predicted to prevail \citep{keres,fumagalli}.
%Therefore, with rare exceptions \citep[e.g.][]{slug,jackpot},
%the community has relied on absorption-line spectroscopy
%to detect and characterize the gas surrounding galaxies
%\citep[e.g.][]{bergeron,steidel10,pro11,tumlinson13}.
%With this experiment, researcher have had recent success
%in characterizing the large-scale flows around star-forming
%galaxies to reveal a net inflow \cite{rakic13} and
%provide unique constraints on the mass of the underlying
%potential well \cite{rakic14}.
%A significant limitation of standard absorption-line
%analysis, especially regarding sutyding galactic flows,
%is the inherent symmetry of the experiment.
%Because one generally lacks any constraint on the distance
%of the gas along the sightline,
%postive or negative velocities with respect to the
%galaxy may be interpreted as gas flowing either towards
%or away from the system.
%So-called `down-the-barrel' observations break this
%symmetry, and have generally provided evidence for flows
%away from galaxies \citep{rupke,martin,weiner,rubin}.
%However, these data are frequently at low spectral resolution
%which limits one's sensitivity to inflowing gas.
%In this paper, we examine the flows of gas in the environments
%of massive galaxies hosting quasars.  Our approach leverages
%a large dataset of quasar pairs \citep{hennawi}
%to use the standard technique of absorption-line spectroscopy
%with background quasars. [word]
%These quasar pairs have angular
%separations that correspond to less than 300\,kpc physical
%separation at the f/g quasar redshift.
%Our previous publications from these quasar pairs
%have established that these galaxies are surrounded
%by a massive, cool, and enriched circumgalactic medium
%\citep{QPQ1,QPQ5,QPQ6,QPQ7}.
%We have collected a sample of ntot~spectra passing within
%300\,kpc from a f/g quasar with a precisely measured
%redshift.  Our primary scientific interests are twofold:
%(i) search for signatures of galactic-scale outflows from the
%central galaxy, presumably driven by recent star-formation
%and/or AGN feedback;
%(ii) characterize the dynamics of galactic flows around these
%massive system.  We further descrbie an aspect of this
%experiment that offers a unique opportunity to study galactic
%flows. [detail??]

We adopt a $\Lambda$CDM cosmology with $\Omega_M=0.26, \Omega_\Lambda=0.74$, and 
$H_0=70\,{\rm km\,s^{-1}\,Mpc^{-1}}$. Distances are proper unless otherwise stated. When referring 
to comoving distances we include explicitly an $h^{-1}$ term and follow modern convention of 
scaling to the Hubble constant $h = H_0/(70\,{\rm km\,s^{-1}\,Mpc^{-1}})$.
%All equivalent width measurements are presented in the rest-frame, unless otherwise specified.

\section{The Experiment}
\label{sec:experiment}

%\input{Tables/tab_sample.tex}
\input{tab_sample.tex}

\begin{figure}
%\includegraphics[width=3in]{Figures/fig_experiment.pdf}
\includegraphics[width=3in]{f01.pdf}
\caption{These panels summarize properties of the QPQ9 dataset. The QPQ survey selects 
quasar pairs of $R_\perp < 300$\,kpc and $z_{\rm fg}>1.6$. Assuming that the foreground quasars 
emit isotropically and at a distance equal to the impact parameter, the enhancement in the UV 
flux relative to the extragalactic UV background, $g_{\rm UV}$, can be estimated. Large 
symbols correspond to foreground quasars with the most precise redshift measurement from 
[\ion{O}{3}]\,5007, while small symbols correspond to $z_{\rm fg}$ 
measurements from \ion{Mg}{2}\,2800, H$\alpha$, or H$\beta$ emission. The top panel shows 
quasar pairs with coverage of \ion{C}{2}\,1334 at $z_{\rm fg}$ in the background quasar spectra. 
The middle panel shows pairs with coverage of \ion{C}{4}\,1548. The bottom panel shows 
pairs with coverage of \ion{Mg}{2}\,2796. 
}
\label{fig:experiment}
\end{figure}

The goal of our experiment is to measure the average velocity fields of the absorption from C$^+$, 
C$^{3+}$, and Mg$^+$ ions associated with the CGM of the galaxies hosting $z\sim 2$ quasars.

From our QPQ survey\footnote{http://www.qpqsurvey.org}, we analyze a subset of systems that 
pass within transverse separation $R_\perp<300$\,kpc from a foreground quasar with 
$z_{\rm fg} > 1.6$. We restrict the sample to foreground quasars with redshift measured from 
\ion{Mg}{2}\,2800, [\ion{O}{3}]\,5007, or H$\alpha$ emission, giving a precision of 
$300\,{\rm km\,s^{-1}}$ or better and an average offset from the systemic redshift of 
$|100\,{\rm km\,s^{-1}}|$ or less. According to \cite{Shen+16}, the [\ion{O}{3}] emission-line 
redshifts have the smallest scatter (intrinsic scatter and measurement error combined) of 
$68\,{\rm km\,s^{-1}}$ about the systemic redshift, and we analyze the sub-sample with 
[\ion{O}{3}] redshifts separately. The [\ion{O}{3}] line has an average blueshift of 
$48\,{\rm km\,s^{-1}}$ about the systemic redshift, which has been added when we compute the 
redshift of the line. The scatter and average offset of [\ion{O}{3}] redshifts reported by 
\cite{Shen+16} is consistent with the numbers reported by \cite{Boroson05} using a larger but 
lower redshift sample. Systemic redshifts measured from \ion{Mg}{2} have a precision of 
$226\,{\rm km\,s^{-1}}$ according to \cite{Shen+16}, and we have taken into account their 
reported median blueshift of $57\,{\rm km\,s^{-1}}$ of \ion{Mg}{2} from the systemic. We note that 
\cite{Richards+02} reported a median redshift of $97\,{\rm km\,s^{-1}}$ of \ion{Mg}{2} from 
[\ion{O}{3}] using a larger but lower redshift sample. In QPQ8, we quantified 
the precision of H$\alpha$ to be $300\,{\rm km\,s^{-1}}$ and the median offset from the systemic 
redshift is close to zero, consistent with the velocity shifts measured by \cite{Shen+11}. 
Although H$\beta$ is a narrow emission-line, we do not consider its redshift sufficiently reliable 
for use as systemic redshift. H$\beta$ redshifts have a large scatter about the systemic 
$\approx400\,{\rm km\,s^{-1}}$, and a large average offset about the systemic 
$\approx100\,{\rm km\,s^{-1}}$ \citep[][QPQ8]{Shen+16}.
Our line-centering algorithm calculates the mode of a line given by
$3\times{\rm median}-2\times{\rm mean}$, applied to the upper 60\% of the emission, while
\cite{Shen+16} calculates the peak of a line. We expect that our line-centering algorithm gives
emission redshifts very comparable to the \cite{Shen+16} algorithm, however. \cite{Shen+16}
states that the difference between the peak and the centroid of an emission-line
is not significant except for the broad line H$\beta$, which we do not use in redshift
measurements. To quantify the above, we further obtain individual measurements of centroids and
peaks in the \cite{Shen+16} sample through private communication. We found there is essentially no
difference between using the centroid versus using the peak for [\ion{O}{3}] emission redshifts,
and there is on average $50\,{\rm km\,s^{-1}}$ difference for \ion{Mg}{2}. We may expect the
difference between the mode and the peak is even smaller. Hence, we argue that the average
systemic bias corrections measured in \cite{Shen+16} may be self-consistently applied to our
measured emission-line redshifts to obtain systemic redshifts.

We further add to the QPQ dataset with quasar pairs selected from the public dataset of 
igmspec\footnote{https://github.com/specdb/igmspec} \citep{Prochaska+17}, which includes the 
spectra from the quasar catalogs based upon the Sloan Digital Sky Survey Seventh Data Release 
\citep{Schneider+10} and the Twelfth Data Release \citep{Paris+17}. We only select pairs with 
$z_{\rm fg}$ measurable using a robust \ion{Mg}{2}\,2800 emission-line. 
We reach a final sample size of 102. 
Figure~\ref{fig:experiment} summarizes the experimental design. We refer the reader to previous 
QPQ publications for the details on the emission-line centering algorithm, data reduction, and
continuum normalization (QPQ1; QPQ6; QPQ8).

As in the previous QPQ papers, we select the quasar pairs to have redshift difference 
$>3000\,{\rm km\,s^{-1}}$, to exclude physically associated binary quasars. The cut on velocity 
difference is motivated by the typical redshift uncertainty of $\approx500\,{\rm km\,s^{-1}}$ of 
the background quasars. In QPQ8, it was required that the observed wavelengths of the 
metal ion transitions fall outside the Ly$\alpha$ forest of the background quasar. In this paper, 
we exclude a small window around the Ly$\alpha$ emission, in additional to the Ly$\alpha$ forest, 
from analysis. For stacked profile analysis, a good estimate of 
the continuum level is necessary. In QPQ8 we found that absorption associated to 
the foreground quasar occurs within $\pm2000\,{\rm km\,s^{-1}}$ around $z_{\rm fg}$. Therefore, it 
is desirable to keep a $\approx\pm3000\,{\rm km\,s^{-1}}$ window relatively free of contamination 
from Ly$\alpha$ forest. Taking into account the redshift uncertainties, 
we decide that at least one transition among \ion{C}{2}\,1334, \ion{C}{4}\,1548, and 
\ion{Mg}{2}\,2796 at $z_{\rm fg}$ must lie redward of 
$(1215.6701+20)\times(1+z_{\rm bg})\,{\rm \AA}$, for a pair to be included in the analysis.

Furthermore, we include only those spectra with average signal-to-noise ratio (S/N) exceeding 5.5 
per rest-frame \AA \ in a $\pm3000\,{\rm km\,s^{-1}}$ window centered on the observed wavelengths 
of the metal ion transitions. This criterion is a compromise between maximizing sample size versus 
maintaining good data quality on the individual sightlines. We find that S/N $>5.5$ per rest-frame 
\AA \ is necessary for properly estimating the continuum, as well as identifying mini-broad 
absorption line systems associated to the background quasar, which will significantly depress the 
flux level. We also require that the region of the spectrum that is $\pm3000\,{\rm km\,s^{-1}}$ 
around a considered metal ion transition does not overlap with strong atmospheric ${\rm O}_2$ 
bands. The ${\rm O}_2$ A- and B-band span 7595\textendash7680\,\AA \ and 6868\textendash6926\,\AA 
\ respectively.

Table~\ref{tab:sample} lists the full QPQ9 sample. In Table~\ref{tab:summary}, we first list the 
sample size, the median $z_{\rm fg}$, and the median $R_\perp$ of the quasar pairs that survive 
the above selection criteria for \ion{C}{2}\,1334, \ion{C}{4}\,1548, and \ion{Mg}{2}\,2796 
respectively. We then provide the summary for the sub-sample with $z_{\rm fg}$ measured from 
[\ion{O}{3}].

\section{Analysis}
\label{sec:analysis}

\subsection{Stacked Profiles}
\label{sec:stacks}

\begin{figure}
%\includegraphics[width=3.5in]{Figures/fig_stacks_and_fits.pdf}
\includegraphics[width=3.5in]{f02.pdf}
\caption{Mean and median absorption centered at \ion{C}{2}\,1334, \ion{C}{4}\,1548, and 
\ion{Mg}{2}\,2796 of the foreground quasars for all QPQ9 pairs. The composites are shown in thin, 
black. Overplotted on the composites are Gaussian fits, normalized to pseudo-continua far away 
from a velocity of $0\,{\rm km\,s^{-1}}$ relative to $z_{\rm fg}$, and are shown in thick, blue. 
For the doublets, a second Gaussian with a fixed mean separation and a tied standard deviation is 
included in the modeling. The absorptions frequently exhibit large velocity widths. The dashed 
lines mark the centroids, which show small positive velocity offsets from $z_{\rm fg}$. The 
$1\sigma$ modeling error for the centroid and dispersion of the \ion{C}{2} mean stack are 
$25\,{\rm km\,s^{-1}}$ and $27\,{\rm km\,s^{-1}}$ respectively, which is also the typical modeling
errors of the other stacks.
}
\label{fig:stacks_and_fits}
\end{figure}

We create composite spectra that average over the intrinsic scatter in quasar environments, 
continuum placement errors, and redshift errors. The individual spectra of background quasars are 
shifted to the rest-frame of the foreground quasars at the transitions of interest. Each spectrum 
has been linearly interpolated onto a fixed velocity grid centered at $z_{\rm fg}$ with bins of 
$100\,{\rm km\,s^{-1}}$. For a velocity bin of this size, it is unnecessary to smooth the data to 
a common spectral resolution. The individual spectra are then combined with a mean or a median 
statistic. Bad pixels in the individual spectra have been masked before generating the 
composites. Since each quasar pair gives an independent probe of the CGM, each pair has an equal 
weighting in the stacked profiles, i.e.\ we do not weight the spectra by the measured S/N near the 
metal ion transitions. Scatter in the stacked spectra is dominated by randomness in the CGM rather 
than scatter in the flux of individual observations. The mean statistic of the individual spectra 
yields a good estimate of the average absorption and preserves equivalent width. The median 
statistic is less sensitive to outliers, however the median opacity at any velocity channel is 
rather small, since the discrete absorbers are spread throughout the entire velocity window. The 
analysis on the median velocity field is thus subject to larger uncertainty. In the following we 
present stacked spectra using both the mean and the median statistic.

In Figure~\ref{fig:stacks_and_fits}, we present mean and median stacks of \ion{C}{2}\,1334, 
\ion{C}{4}\,1548, and \ion{Mg}{2}\,2796 absorption of the QPQ9 sample. We focus on the 
analysis results of the \ion{C}{2} mean stack. \ion{C}{4} and \ion{Mg}{2} are doublet transitions 
and it is more challenging to analyze their kinematics.
%To assess the average absorption of these transitions,
%we have combined the data with a novel technique.
%We have fitted the combined dataset using a cubic b-spline
%with knots at every 100\,kms.  We weighted each data
%point by the velocity width of the pixel to insure that 
%each sightline makes an equal contribution to the stack.
%These b-spline profiles are presented in the figure
%and a trivial integration gives the reported equivalent
%widths.
%[consider down-weighting systems with $<3\sigma$ EW]
%[Comnent on EW from 'standard' technique]
Two results are evident in Figure~\ref{fig:stacks_and_fits}: (i) the mean \ion{C}{2} stack 
exhibits excess absorption spanning a large velocity width; (ii) the mean absorption is likely 
skewed toward positive velocities. We have 40 pairs in the QPQ9 sample for \ion{C}{2} analysis
and an even larger number of pairs for \ion{C}{4} and \ion{Mg}{2} analyses, and the total number
of individual components can only be larger than the number of absorption systems. Hence, although
QPQ8 showed the distribution of absorbing components in velocity is non-Gaussian for each
individual quasar-host associated system, the central limit theorem applies and the distribution
of all absorbing components should tend toward Gaussian distribution. A visual inspection of the 
stack confirms that a broader absorption component is not apparent. Therefore, to model the 
absorption, we may introduce a Gaussian profile while allowing a constant pseudo-continuum level 
to vary. We perform $\chi^2$ minimization with each channel given equal weight. From the best-fit 
to the data, we measure the centroid of the \ion{C}{2} stack to be $+232\,{\rm km\,s^{-1}}$ and
the $1\sigma$ dispersion to be $293\,{\rm km\,s^{-1}}$. The centroid suggests an asymmetry that
contradicts the standard expectation, while the dispersion suggests extreme kinematics. The median 
stack, on the other hand, shows weaker absorption, and the Gaussian model has a centroid with 
smaller offset and comparable dispersion.

We also create mean and median stack for the sub-sample with [\ion{O}{3}] redshifts and model the 
absorption with Gaussian best-fit. The \ion{C}{2} mean stack for this sub-sample has a centroid at 
$+235\,{\rm km\,s^{-1}}$, and a dispersion of $330\,{\rm km\,s^{-1}}$, consistent with the full
sample.

To model the mean and median absorption of \ion{C}{4}\,1548 and \ion{Mg}{2}\,1796, we introduce a  
second Gaussian with separation equal to the doublet separation ($498\,{\rm km\,s^{-1}}$ and 
$769\,{\rm km\,s^{-1}}$ respectively), and tie the dispersion of the two lines in a doublet. We
allow the doublet ratio to vary from 2:1 to 1:1. The modeling results show that the velocity fields
of \ion{C}{4} and \ion{Mg}{2} are consistent with \ion{C}{2}, i.e.\ large dispersion and centroid
is skewed toward positive velocities.

The above analyses are summarized in Table~\ref{tab:summary}. The Gaussian models normalized to 
pseudo-continuum are overplotted on the data stacks in Figure~\ref{fig:stacks_and_fits}.

%\input{Tables/tab_summary.tex}
\input{tab_summary.tex}

\subsection{Interpretation of the asymmetric absorption}
\label{sec:significance_+ve}

Given the small sample size, one must scrutinize the statistical significance of the measured 
velocity offsets. To assess the statistical significance, we perform a boostrap analysis by 
randomly resampling from the full sample 10000 times. We introduce a Gaussian absorption profile 
to model each bootstrap realization, and quote the standard deviation in the bootstrap 
realizations to be the scatter in the centroids of the data. The scatters 
$\approx100\,{\rm km\,s^{-1}}$ are comparable to the measured offsets, indicating large intrinsic 
variation in quasar CGM environments (see also QPQ8). In Figure~\ref{fig:histogram_cen}, we show 
the distribution of the absorption centroids from bootstrapping on the \ion{C}{2} mean stack. We 
find that 97\% of the centroids are at positive velocities. We place a generous $3\sigma$ upper
limit to the small offset of the centroid from $z_{\rm fg}$ for the \ion{C}{2} mean absorption at 
$\delta v<+526\,{\rm km\,s^{-1}}$. Given the large intrinsic scatter, we do not attempt to explore
whether there exists relative asymmetry among the \ion{C}{2}, \ion{C}{4}, and \ion{Mg}{2} 
absorption.

One may also ask whether the measured positive offsets come from systematic bias in redshift 
measurements due to the Baldwin effect \citep{Baldwin77}. \cite{Shen16} reported that, the 
[\ion{O}{3}] emission of $z\sim2$ quasars is more asymmetric and weaker than that in typically 
less luminous low-$z$ quasars. To test for this potential source of bias, we create another mean 
stack at \ion{C}{2}\,1334 by replacing the [\ion{O}{3}] redshifts by a redshift measured from the
more symmetric \ion{Mg}{2} or H$\alpha$ emission when available. We are able to replace for 11 out
of the 15 systems with [\ion{O}{3}] redshifts in the original sample. The new stack is similiar in
velocity structure and again shows a positive offset $\approx+303\,{\rm km\,s^{-1}}$. We thus
conclude that our algorithm for measuring redshifts is not severely biased by the blue wing of the
[\ion{O}{3}] emission-line.

Motivated by the study of \ion{Mg}{2} absorbers surrounding $z\sim1$ quasars by \cite{Johnson+15}, 
we also generate a mean-stacked spectrum for \ion{Mg}{2}\,2796 for lower redshift quasar pairs.  
We select quasars with $0.4<z_{\rm fg}<1.6$ and use the same other selection criteria as the main 
QPQ9 sample. The quasar pairs are selected from the igmspec database, with $z_{\rm fg}$ measured 
by \cite{HewettWild10}. For quasars with $z_{\rm fg}<0.84$, as redshift determination is dominated 
by [\ion{O}{3}] emission, a shift of $+48\,{\rm km\,s^{-1}}$ is applied to bring the emission-line 
redshift to the systemic. For quasars with $0.84<z_{\rm fg}<1.6$, as redshift determination is 
dominated by \ion{Mg}{2} emission, a shift of $+57\,{\rm km\,s^{-1}}$ is applied. There are 233
pairs selected, with a median $z_{\rm fg}$ of 0.90 and a median $R_\perp$ of 208\,kpc. We present 
the mean stack in Figure~\ref{fig:stack_z1}. The absorption is weaker than the $z\sim2$ main QPQ9 
sample. Gaussian absorption models fitted to the stack recover a centroid of 
$-11\pm379\,{\rm km\,s^{-1}}$ and a dispersion of $172\,{\rm km\,s^{-1}}$. The average offset
from $0\,{\rm km\,s^{-1}}$ is much smaller than the offsets in the $z\sim2$ sample.

Since the large scatter in the centroids represents intrinsic variation rather than redshift 
errors, and the \ion{Mg}{2} stack for lower redshift suggests a different centroid, we 
consider that the asymmetry signal in the $z\sim2$ sample is likely to be real. In the Discussion 
section, we discuss two possible explanations for the asymmetry.

\begin{figure}
%\includegraphics[width=2.5in]{Figures/fig_histogram_cen.pdf}
\includegraphics[width=2.5in]{f03.pdf}
\caption{Histogram of the absorption centroids of 10000 bootstrap realizations of the data sample 
for the \ion{C}{2} mean stack. 97\% of the centroids are positively offset from $z_{\rm fg}$.}
\label{fig:histogram_cen}
\end{figure}

\begin{figure}
%\includegraphics[width=2.5in]{Figures/fig_stack_z1.pdf}
\includegraphics[width=2.5in]{f04.pdf}
\caption{Mean stack at \ion{Mg}{2}\,2796, for a lower redshift sample of quasar pairs at 
$z\sim0.9$. Line-style coding is the same as in Figure~\ref{fig:stacks_and_fits}. The centroid is 
approximately at $0\,{\rm \,km\,s^{-1}}$, and the absorption is weaker than the main QPQ9 sample. 
}
\label{fig:stack_z1}
\end{figure}

\subsection{Interpretation of the large velocity fields}
\label{sec:significance_width}

\begin{figure}
%\includegraphics[width=2.5in]{Figures/fig_monte.pdf}
\includegraphics[width=2.5in]{f05.pdf}
\caption{The same \ion{C}{2} mean stack shown in the first panel of 
Figure~\ref{fig:stacks_and_fits} is shown in thin, black. In thick, green, we overplot the 
Gaussian absorption model of the Monte Carlo simulations generated from a purely clustering 
argument. The model from clustering analysis is multiplied to the pseudo-continuum level of the 
stack of the observational data, shifted to the centroid of the stack of the data, and broadened 
by the mean redshift error in the data. The model is a good description of the data, with a 
dispersion within modeling error of the dispersion in the data. We overplot in dashed, yellow an 
absorption profile with $\sigma_v=554\,{\rm km\,s^{-1}}$, which is larger than the observed width
by three times the standard deviation in the bootstrap analysis. Motions that produce a velocity 
width larger than this can be ruled out. We overplot in dashed, magenta an absorption profile with 
$\sigma_v=214\,{\rm km\,s^{-1}}$, which is smaller than the observed width by three times the
modeling error. Unless gravitational and Hubble flows together with redshift error broadening 
produce a velocity width smaller than this, extra dynamical processes (e.g.\ outflows) will not be 
required to explain the observation. 
}
\label{fig:monte}
\end{figure}

\begin{figure}
%\includegraphics[width=2.5in]{Figures/fig_contour.pdf}
\includegraphics[width=2.5in]{f06.pdf}
\caption{
Probability distributions of the parameters $W_{\rm CII}$ and $M_{\rm halo}$. The plot shows the 
degeneracy between $W_{\rm CII}$ and $M_{\rm halo}$ in recovering the intrinsic width of the 
absorption profile. We mark contours for points that produce an absorption 
profile of width that is 1, 2, and 3 times the modeling error away from the observed intrinsic 
width. The intrinsic velocity width corresponding to typical QPQ halo mass is marked with a plus 
sign, and is well contained within the 1$\sigma$ contour.}
\label{fig:contour}
\end{figure}

Under the assumption that the intrinsic dispersion and the redshift uncertainty add in 
quadrature to give the observed width, we solve for the intrinsic dispersion in the \ion{C}{2} 
mean stack. For the full QPQ9 sample, with the mean 
$\sigma_{{\rm error}(z)}^{\rm full}=189\,{\rm km\,s^{-1}}$, we recover
$\sigma_{\rm intrinsic}^{\rm full}=224\,{\rm km\,s^{-1}}$. For the sub-sample with [\ion{O}{3}]
redshifts, we recover $\sigma_{\rm intrinsic}^{\rm [OIII]}=323\,{\rm km\,s^{-1}}$.

\cite{Eftekharzadeh+15} measured the clustering of quasars in the range $2.2<z<2.8$ while 
\cite{RodriguezTorres+17} measured the clustering of quasars in the range $1.8<z<2.2$. They 
estimated that these quasars are hosted by dark matter halos with mass 
$M_{\rm halo}=10^{12.5}\,M_\odot$ and $M_{\rm halo}=10^{12.6}\,M_\odot$ respectively. If dark
matter halos hosting QPQ9 quasars have a characteristic mass $10^{12.6}\,M_\odot$ and follow an 
NFW profile \citep{NavarroFrenkWhite97} with concentration parameter $c=4$, at $z\approx2$ the
maximum circular velocity is $345\,{\rm km\,s^{-1}}$. \cite{TormenBouchetWhite97} found that the
maximum circular velocity is $\approx1.4$ times the maximum of the average one-dimensional
root-mean-square velocity $\sigma_{\rm rms}$. Hence, the average line-of-sight rms velocity
typical of QPQ9 halos is $\sigma_{\rm rms}=246\,{\rm km\,s^{-1}}$. In QPQ8, we estimated the
probability of intercepting a random optically thick absorber is 4\%, and clustering would only
increase that to 24\%. Although motions due to Hubble flows do not dominate, they nevertheless
contribute to the observed dispersion. We can investigate
whether gravitational motions and Hubble flows are sufficient to reproduce the dispersion in the
data, using Monte Carlo methods to simulate the absorption signals.

Since \ion{C}{2} systems arise in optically thick absorbers, we may adopt the clustering analysis 
results of QPQ6. In the absence of clustering, the expected number of absorbers per unit redshift 
interval for Lyman limit systems, super Lyman limit systems, and damped Ly$\alpha$ systems are 
respectively $\ell_{\rm IGM}^{\rm LLS}(z)\approx1.05((1+z)/(1+2.5))^{2.1}$, 
$\ell_{\rm IGM}^{\rm SLLS}(z)\approx0.44((1+z)/(1+2.5))^{2.1}$, and 
$\ell_{\rm IGM}^{\rm DLA}(z)\approx0.2((1+z)/(1+2.5))^{2.1}$. The quasar-absorber correlation 
functions for Lyman limit systems, super Lyman limit systems, and damped Ly$\alpha$ systems are 
respectively $\xi_{\rm QA}^{\rm LLS}(r)=(r/(12.5\,h^{-1}\,{\rm Mpc}))^{-1.68}$, 
$\xi_{\rm QA}^{\rm SLLS}(r)=(r/(14.0\,h^{-1}\,{\rm Mpc}))^{-1.68}$, and 
$\xi_{\rm QA}^{\rm DLA}(r)=(r/(3.9\,h^{-1}\,{\rm Mpc}))^{-1.6}$. 
For each quasar pair, we calculate the expected number of optically thick absorbers within 
$\pm3000\,{\rm km\,s^{-1}}$ at a distance $R_\perp$ from the foreground quasar and at 
$z_{\rm fg}$. Then we generate 1000 mock sightlines. The number of absorbers for each mock 
spectrum is randomly selected from a Poisson distribution with mean equal to the expected number 
calculated as above. The absorbers are randomly assigned Hubble velocities, with a probability 
distribution according to the quasar-absorber correlation functions. The absorbers are randomly 
assigned additional peculiar velocities drawn from a normal distribution with mean equal to 
$0\,{\rm km\,s^{-1}}$ and scatter equal to $\sigma_{\rm rms}$. For each absorber, we assume a rest
equivalent width for \ion{C}{2} $W_{\rm CII}$ and a Gaussian absorption profile. We repeat the 
above procedure for all 33 quasar pairs, and create a mean stack of the 40000 mock spectra
generated. We fit a Gaussian absorption profile multiplied to a constant continuum level to model 
the stack of mock spectra. We adjust the $W_{\rm CII}$ adopted for the absorbers until the 
amplitude of the best-fit Gaussian of the stack of mock spectra matches the amplitude of the stack
of the observational data. We find that $W_{\rm CII}\approx0.5\,{\rm \AA}$ well reproduces the
amplitude, and the dispersion of the Gaussian absorption model is insensitive to the assumed
$W_{\rm CII}$ or line profile for one absorber.

In Figure~\ref{fig:monte}, we show a comparison of the observational data stack and the Gaussian 
absorption model of the Monte Carlo simulations. In the figure, the Gaussian absorption model is 
broadened by the mean redshift error by adding it in quadrature to the dispersion in the model. 
The resulting stack of mock spectra has a $1\sigma$ dispersion of $282\,{\rm km\,s^{-1}}$. This is
about 2 times the modeling error away from the intrinsic dispersion in the \ion{C}{2} mean stack 
for the full sample, and about 2 times the modeling error away from the intrinsic dispersion in the
stack of the sub-sample with [\ion{O}{3}] redshifts as well. We also test for the
sensitivity of this measured dispersion to the correlation functions adopted. The QPQ6 clustering 
analysis is performed on only the strongest absorber near $z_{\rm fg}$, and a low $R_\perp$ 
sightline may in fact intercept more than one optically thick absorber. We double the number of 
absorbers for each mock sightline, and find the measured dispersion only increases by several 
${\rm km\,s^{-1}}$. Thus, the hypothesis that the observed velocity width is only produced by a 
combination of gravitational motions and Hubble flows cannot be ruled out. Extra dynamical 
processes are not necessary to explain the large velocity fields.

From the bootstrap analysis of the observed dispersion, a width of
$\sigma_v>554\,{\rm km\,s^{-1}}$ would be larger than the observed by three times the scatter
(overplotted in Figure~\ref{fig:monte}). In other words, taking into account the redshift errors,
motions in addition to gravitational and Hubble flows that will produce an intrinsic dispersion
$\sigma_v>521\,{\rm km\,s^{-1}}$ can be ruled out.

On the contrary, 
$\sigma_{\rm rms}<70\,{\rm km\,s^{-1}}$ together with Hubble velocities and
broadening by redshift errors will result in a velocity width that is more than three times the 
modeling error away from the observed width (overplotted in Figure~\ref{fig:monte}). This implies 
that, unless the characteristic $M_{\rm halo}<10^{11.0}\,M_\odot$,
additional dynamical processes are not required to explain the observed width.

In Figure~\ref{fig:contour}, We show the probability distributions of the degenerate parameters 
$W_{\rm CII}$ and $M_{\rm halo}$ in recovering the intrinsic width of the absorption profile. We 
require that the amplitude of the absorption is reproduced within 3 times its modeling error, and 
mark contours for points in ($W_{\rm CII}$, $M_{\rm halo}$)-space that produce an absorption 
profile of width within 1, 2, and 3 times the modeling error of the observed intrinsic width. 
From the figure, a higher $W_{\rm CII}$ means the $M_{\rm halo}$ that will reproduce the observed 
width is higher. If there are no extra dynamical processes, the intrinsic velocity width 
corresponding to typical QPQ halo mass is contained within the 2$\sigma$ contour.

\section{Discussion}
\label{sec:discussion}

\begin{figure*}[!th]
\begin{minipage}[!bp]{0.28\textwidth}
%\includegraphics[width=\textwidth]{Figures/fig_monopolar.pdf}
\includegraphics[width=\textwidth]{f08.pdf}
\caption{A cartoon showing a unipolar quasar. The gas observed in low- to intermediate-ion 
absorption preferentially lies behind the quasar, and is shadowed from the ionizing radiation. 
}
\label{fig:monopolar}
\end{minipage}%
\hspace{0.3in}
\begin{minipage}[!bp]{0.67\textwidth}
%\includegraphics[width=\textwidth]{Figures/fig_lighttravel.pdf}
\includegraphics[width=\textwidth]{f09.pdf}
\caption{A cartoon showing the finite lifetime of quasar episodes as an explanation to the 
asymmetric absorption. The setup on the left shows that the foreground quasar has not been shining 
long enough for its ionizing radiation to reach the gas behind it, when the light from the 
background quasar reaches. The setup on the right shows the scenario after an amount of time 
comparable to the light travelling time across CGM scale. Gas in front of the foreground quasar 
has been ionized, by the time the light from the background quasar reaches. 
}
\label{fig:lighttravel}
\end{minipage}
\end{figure*}

At low redshift, both symmetric and asymmetric ionization cones around quasars and AGNs are
observed. In the extended emission-line region of 4C37.43, most of the [\ion{O}{3}] emission is
blueshifted \citep{FuStockton07}, but there are counter examples in less extended sources in
\cite{FuStockton09}. Recently, there has been Fabry-Perot interferometric data for less extended
narrow-line regions in more nearby sources \citep{Keel+15,Keel+17}, which show mostly symmetric
velocity fields and gas distributions.

In the following, we explore two possible explanations for the putative non-dynamical process that
provides the asymmetry.

One possibility is an asymmetric radiation field that preferentially ionizes the gas moving toward 
the observer, where the quasar is known to shine. Alternatively, the asymmetric radiation field 
may preferentially ionizes the gas at smaller Hubble velocity than the quasar. In 
Figure~\ref{fig:monopolar}, we show a cartoon of a quasar that is blocked in the direction 
pointing away from the observer. The gas observed in absorption preferentially lies behind the 
quasar. The asymmetric absorption arises from a transverse proxmity effect. 
\cite{Roos+15} and \cite{GaborBournaud14} performed simulations of a high-redshift disk galaxy 
including thermal AGN feedback and calculated radiative transfer in post-processing. They found 
the ionization radiation is typically asymmetric, due to either a dense clump that lies on one 
edge of the black hole or the black hole's location being slightly above the disk.
\cite{FaucherGiguere+16} represents the only simulation so far that is able to reproduce the
substantial amount of cool gas in quasar-mass haloes. We eagerly await their group to compare the
fraction of gas in inflows and outflows.

As a test of reasonableness, we calculate that, for a quasar opening angle of $120^\circ$ and
keeping $\sigma_{\rm rms}$ at the expected value for the typical QPQ9 halo, our Monte Carlo
simulations may reproduce the observed centroid and intrinsic dispersion of the \ion{C}{2} mean
absorption to within the modeling error.

Another possible explanation arises from the finite lifetime of quasar episodes. 
Figure~\ref{fig:lighttravel} presents a cartoon for a light travel time argument. Light from the 
background quasar will travel smaller distance to reach the gas behind the foreground quasar 
than the gas in front of it. The light from the background quasar may arrive at the gas behind 
the foreground quasar before the ionizing radiation from the foreground quasar arrives. In 
QPQ7 and QPQ8, we reported that the enhancement in metal ion absorption relative to IGM average 
extends to at least $\approx200$\,kpc. This distance corresponds to a light travelling time of 
$\sim10^6$\,yr, which lies within existing constraints of quasar lifetime from observations 
\citep[e.g.][]{Martini04} and simulations \citep[e.g.][]{Hopkins+05}.

%As a test of reasonableness, we calculate that, at $R_\perp\approx50\textendash150$\,kpc, radial
%speeds relative to the foreground quasar of $500\textendash700\,{\rm km\,s^{-1}}$ and quasar
%opening angles of $120^\circ\textendash140^\circ$ may reproduce the observed centroid and
%intrinsic dispersion of the \ion{C}{2} mean absorption.
The first explanation above to asymmetric
absorption requires that the quasars emit their ionizing radiation anisotropically, while the 
second explanation requires the quasars emit their ionizing radiation intermittently. Both 
explanations will require the line-of-sight motions of the gas to be in either outflow or Hubble 
flow. Were the gas flowing into the galaxy instead, the velocity centroid would be negative.
Under both scenarios, lows ions should be redshifted while high ions should be blueshifted.
Using Cloud photoionization models, we find that, for the typical quasar luminosity of our sample,
a CGM gas that is directly illuminated is highly ionized and shows only marginally detectable
\ion{C}{4} absorption. The redshifted \ion{C}{4} may be regarded as an intermediate, and a
blueshifted absorption signal needs to be searched in a higher ion.

Motivated by the asymmetry found in metal ion absorption in the CGM using precise $z_{\rm fg}$ 
measurements, and the asymmetry found by \cite{KirkmanTytler08} in \ion{H}{1} on larger 
scales, we are assembling a sample of quasar pairs with precise $z_{\rm fg}$ measurements to study 
this asymmetry in \ion{H}{1} (J. F. Hennawi et al. 2017, in preparation). In conclusion, we 
observe large and positively skewed velocity fields in absorption, of metal ions 
in the CGM of $z\sim2$ massive galaxies hosting quasars. We argue that the velocity fields can be 
explained if the detected gas is in gravitational motion, Hubble flow, or outflow, and the quasars 
either shine preferentially toward the observer or shine intermittently.

\acknowledgements

JXP and MWL acknowledge support from the National Science Foundation (NSF) grants AST-1010004 and 
AST-1412981. The authors gratefully acknowledge the support which enabled these observations at 
the Keck, Gemini, Large Binocular Telescope, Very Large Telescope, Las Campanas, and Palomar 
Observatories. The authors acknowledge the sharing of private data by Yue Shen. MWL thanks Hai Fu
for a discussion on extended emission-line regions, T.-K. Chan for a discussion on CGM
simulations, and Chi Po Choi for a discussion on statistics.

\appendix

\section{Line-of-sight absorption}
\label{sec:appendix}

\begin{figure}
%\includegraphics[width=3.4in]{Figures/fig_stacks_fg.pdf}
\includegraphics[width=3.4in]{f07.pdf}
\caption{Mean stacks of the foreground quasar spectra at \ion{C}{2}\,1334, \ion{C}{4}\,1548, and
\ion{Mg}{2}\,2796 for the QPQ9 sample. For \ion{C}{2} and \ion{Mg}{2}, the mean absorption is
weaker than that in the background stacks, and there is no evidence for an excess at $z_{\rm fg}$.
The stack for \ion{C}{4}, which includes line-of-sight absorbers at all distances, shows a large,
blueshifted mean velocity field.
}
\label{fig:stacks_fg}
\end{figure}

In the previous QPQ papers, we argued that optically thick absorbers in the vicinity of quasars
are distributed anisotropically. We now have the means to show this anisotropic clustering
explicitly. Given that the techniques for stacking spectra are established, it is straightforward
to apply the same techniques to stack the foreground quasar spectra. In
Figure~\ref{fig:stacks_fg}, we present mean stacks of foreground quasar spectra for
the QPQ9 sample. We require that the spectra survive a S/N cut of 5.5 per rest-frame \AA\ at
\ion{C}{2}\,1334, \ion{C}{4}\,1548, or \ion{Mg}{2}\,2796 at $z_{\rm fg}$. In constrast to the
large equivalent widths exhibited in the stacks of background spectra, \ion{C}{2} and \ion{Mg}{2}
mean absorption along the line-of-sight to the foreground quasars is weaker, and an excess at
$z_{\rm fg}$ is absent. This supports a scenario where the ionizing radiation of the foreground
quasars are anisotropic and/or intermittent. For \ion{C}{4}\,1548, this stack of all line-of-sight
absorbers, which include absorbers intrinsic to and far away from the quasar, shows a large,
blueshifted velocity field. This excess \ion{C}{4} absorption has been well studied as narrow
associated absorption line systems \citep[e.g., ][]{Wild+08}.

%\bibliographystyle{/Users/lwymarie/Documents/Bibli/apj}
%\bibliography{/Users/lwymarie/Documents/Bibli/allrefs}

\begin{thebibliography}{52}
\expandafter\ifx\csname natexlab\endcsname\relax\def\natexlab#1{#1}\fi

\bibitem[{{Baldwin}(1977)}]{Baldwin77}
{Baldwin}, J.~A. 1977, \apj, 214, 679

\bibitem[{{Bergeron} \& {Boisse}(1991)}]{BergeronBoisse91}
{Bergeron}, J., \& {Boisse}, P. 1991, Advances in Space Research, 11, 241

\bibitem[{{Boroson}(2005)}]{Boroson05}
{Boroson}, T. 2005, \aj, 130, 381

\bibitem[{{Cantalupo} {et~al.}(2014){Cantalupo}, {Arrigoni-Battaia},
  {Prochaska}, {Hennawi}, \& {Madau}}]{Cantalupo+14}
{Cantalupo}, S., {Arrigoni-Battaia}, F., {Prochaska}, J.~X., {Hennawi}, J.~F.,
  \& {Madau}, P. 2014, \nat, 506, 63

\bibitem[{{Churchill} {et~al.}(2012){Churchill}, {Kacprzak}, {Steidel},
  {Spitler}, {Holtzman}, {Nielsen}, \& {Trujillo-Gomez}}]{Churchill+12}
{Churchill}, C.~W., {Kacprzak}, G.~G., {Steidel}, C.~C., {Spitler}, L.~R.,
  {Holtzman}, J., {Nielsen}, N.~M., \& {Trujillo-Gomez}, S. 2012, \apj, 760, 68

\bibitem[{{Eftekharzadeh} {et~al.}(2015){Eftekharzadeh}, {Myers}, {White},
  {Weinberg}, {Schneider}, {Shen}, {Font-Ribera}, {Ross}, {Paris}, \&
  {Streblyanska}}]{Eftekharzadeh+15}
{Eftekharzadeh}, S., {Myers}, A.~D., {White}, M., {Weinberg}, D.~H.,
  {Schneider}, D.~P., {Shen}, Y., {Font-Ribera}, A., {Ross}, N.~P., {Paris},
  I., \& {Streblyanska}, A. 2015, \mnras, 453, 2779

\bibitem[{{Faucher-Gigu{\`e}re} {et~al.}(2016){Faucher-Gigu{\`e}re},
  {Feldmann}, {Quataert}, {Kere{\v s}}, {Hopkins}, \&
  {Murray}}]{FaucherGiguere+16}
{Faucher-Gigu{\`e}re}, C.-A., {Feldmann}, R., {Quataert}, E., {Kere{\v s}}, D.,
  {Hopkins}, P.~F., \& {Murray}, N. 2016, \mnras, 461, L32

\bibitem[{{Fu} \& {Stockton}(2007)}]{FuStockton07}
{Fu}, H., \& {Stockton}, A. 2007, \apj, 666, 794

\bibitem[{{Fu} \& {Stockton}(2009)}]{FuStockton09}
---. 2009, \apj, 690, 953

\bibitem[{{Fumagalli} {et~al.}(2011){Fumagalli}, {Prochaska}, {Kasen}, {Dekel},
  {Ceverino}, \& {Primack}}]{Fumagalli+11}
{Fumagalli}, M., {Prochaska}, J.~X., {Kasen}, D., {Dekel}, A., {Ceverino}, D.,
  \& {Primack}, J.~R. 2011, \mnras, 418, 1796

\bibitem[{{Gabor} \& {Bournaud}(2014)}]{GaborBournaud14}
{Gabor}, J.~M., \& {Bournaud}, F. 2014, \mnras, 441, 1615

\bibitem[{{Gauthier}(2013)}]{Gauthier13}
{Gauthier}, J.-R. 2013, \mnras, 432, 1444

\bibitem[{{Hennawi} {et~al.}(2006){Hennawi}, {Prochaska}, {Burles}, {Strauss},
  {Richards}, {Schlegel}, {Fan}, {Schneider}, {Zakamska}, {Oguri}, {Gunn},
  {Lupton}, \& {Brinkmann}}]{QPQ1}
{Hennawi}, J.~F., {Prochaska}, J.~X., {Burles}, S., {Strauss}, M.~A.,
  {Richards}, G.~T., {Schlegel}, D.~J., {Fan}, X., {Schneider}, D.~P.,
  {Zakamska}, N.~L., {Oguri}, M., {Gunn}, J.~E., {Lupton}, R.~H., \&
  {Brinkmann}, J. 2006, \apj, 651, 61

\bibitem[{{Hennawi} {et~al.}(2015){Hennawi}, {Prochaska}, {Cantalupo}, \&
  {Arrigoni-Battaia}}]{Hennawi+15}
{Hennawi}, J.~F., {Prochaska}, J.~X., {Cantalupo}, S., \& {Arrigoni-Battaia},
  F. 2015, Science, 348, 779

\bibitem[{{Hewett} \& {Wild}(2010)}]{HewettWild10}
{Hewett}, P.~C., \& {Wild}, V. 2010, \mnras, 405, 2302

\bibitem[{{Ho} {et~al.}(2017){Ho}, {Martin}, {Kacprzak}, \&
  {Churchill}}]{Ho+17}
{Ho}, S.~H., {Martin}, C.~L., {Kacprzak}, G.~G., \& {Churchill}, C.~W. 2017,
  \apj, 835, 267

\bibitem[{{Hopkins} {et~al.}(2005){Hopkins}, {Hernquist}, {Martini}, {Cox},
  {Robertson}, {Di Matteo}, \& {Springel}}]{Hopkins+05}
{Hopkins}, P.~F., {Hernquist}, L., {Martini}, P., {Cox}, T.~J., {Robertson},
  B., {Di Matteo}, T., \& {Springel}, V. 2005, \apjl, 625, L71

\bibitem[{{Johnson} {et~al.}(2015){Johnson}, {Chen}, \&
  {Mulchaey}}]{Johnson+15}
{Johnson}, S.~D., {Chen}, H.-W., \& {Mulchaey}, J.~S. 2015, \mnras, 452, 2553

\bibitem[{{Keel} {et~al.}(2017){Keel}, {Lintott}, {Maksym}, {Bennert},
  {Chojnowski}, {Moiseev}, {Smirnova}, {Schawinski}, {Sartori}, {Urry},
  {Pancoast}, {Schirmer}, {Scott}, {Showley}, \& {Flatland}}]{Keel+17}
{Keel}, W.~C., {Lintott}, C.~J., {Maksym}, W.~P., {Bennert}, V.~N.,
  {Chojnowski}, S.~D., {Moiseev}, A., {Smirnova}, A., {Schawinski}, K.,
  {Sartori}, L.~F., {Urry}, C.~M., {Pancoast}, A., {Schirmer}, M., {Scott}, B.,
  {Showley}, C., \& {Flatland}, K. 2017, \apj, 835, 256

\bibitem[{{Keel} {et~al.}(2015){Keel}, {Maksym}, {Bennert}, {Lintott},
  {Chojnowski}, {Moiseev}, {Smirnova}, {Schawinski}, {Urry}, {Evans},
  {Pancoast}, {Scott}, {Showley}, \& {Flatland}}]{Keel+15}
{Keel}, W.~C., {Maksym}, W.~P., {Bennert}, V.~N., {Lintott}, C.~J.,
  {Chojnowski}, S.~D., {Moiseev}, A., {Smirnova}, A., {Schawinski}, K., {Urry},
  C.~M., {Evans}, D.~A., {Pancoast}, A., {Scott}, B., {Showley}, C., \&
  {Flatland}, K. 2015, \aj, 149, 155

\bibitem[{{Kere\v{s}} {et~al.}(2009){Kere\v{s}}, {Katz}, {Fardal}, {Dav{\'e}},
  \& {Weinberg}}]{Keres+09}
{Kere\v{s}}, D., {Katz}, N., {Fardal}, M., {Dav{\'e}}, R., \& {Weinberg}, D.~H.
  2009, \mnras, 395, 160

\bibitem[{{Kirkman} \& {Tytler}(2008)}]{KirkmanTytler08}
{Kirkman}, D., \& {Tytler}, D. 2008, \mnras, 391, 1457

\bibitem[{{Lau} {et~al.}(2016){Lau}, {Prochaska}, \& {Hennawi}}]{QPQ8}
{Lau}, M.~W., {Prochaska}, J.~X., \& {Hennawi}, J.~F. 2016, \apjs, 226, 25

\bibitem[{{Martin}(2005)}]{Martin05}
{Martin}, C.~L. 2005, \apj, 621, 227

\bibitem[{{Martini}(2004)}]{Martini04}
{Martini}, P. 2004, Coevolution of Black Holes and Galaxies, 169

\bibitem[{{Muzahid} {et~al.}(2015){Muzahid}, {Kacprzak}, {Churchill},
  {Charlton}, {Nielsen}, {Mathes}, \& {Trujillo-Gomez}}]{Muzahid+15}
{Muzahid}, S., {Kacprzak}, G.~G., {Churchill}, C.~W., {Charlton}, J.~C.,
  {Nielsen}, N.~M., {Mathes}, N.~L., \& {Trujillo-Gomez}, S. 2015, \apj, 811,
  132

\bibitem[{{Navarro} {et~al.}(1997){Navarro}, {Frenk}, \&
  {White}}]{NavarroFrenkWhite97}
{Navarro}, J.~F., {Frenk}, C.~S., \& {White}, S.~D.~M. 1997, \apj, 490, 493

\bibitem[{{P{\^a}ris} {et~al.}(2017){P{\^a}ris}, {Petitjean}, {Ross}, {Myers},
  {Aubourg}, {Streblyanska}, {Bailey}, {Armengaud}, {Palanque-Delabrouille},
  {Y{\`e}che}, {Hamann}, {Strauss}, {Albareti}, {Bovy}, {Bizyaev}, {Niel
  Brandt}, {Brusa}, {Buchner}, {Comparat}, {Croft}, {Dwelly}, {Fan},
  {Font-Ribera}, {Ge}, {Georgakakis}, {Hall}, {Jiang}, {Kinemuchi},
  {Malanushenko}, {Malanushenko}, {McMahon}, {Menzel}, {Merloni}, {Nandra},
  {Noterdaeme}, {Oravetz}, {Pan}, {Pieri}, {Prada}, {Salvato}, {Schlegel},
  {Schneider}, {Simmons}, {Viel}, {Weinberg}, \& {Zhu}}]{Paris+17}
{P{\^a}ris}, I., {Petitjean}, P., {Ross}, N.~P., {Myers}, A.~D., {Aubourg},
  {\'E}., {Streblyanska}, A., {Bailey}, S., {Armengaud}, {\'E}.,
  {Palanque-Delabrouille}, N., {Y{\`e}che}, C., {Hamann}, F., {Strauss}, M.~A.,
  {Albareti}, F.~D., {Bovy}, J., {Bizyaev}, D., {Niel Brandt}, W., {Brusa}, M.,
  {Buchner}, J., {Comparat}, J., {Croft}, R.~A.~C., {Dwelly}, T., {Fan}, X.,
  {Font-Ribera}, A., {Ge}, J., {Georgakakis}, A., {Hall}, P.~B., {Jiang}, L.,
  {Kinemuchi}, K., {Malanushenko}, E., {Malanushenko}, V., {McMahon}, R.~G.,
  {Menzel}, M.-L., {Merloni}, A., {Nandra}, K., {Noterdaeme}, P., {Oravetz},
  D., {Pan}, K., {Pieri}, M.~M., {Prada}, F., {Salvato}, M., {Schlegel}, D.~J.,
  {Schneider}, D.~P., {Simmons}, A., {Viel}, M., {Weinberg}, D.~H., \& {Zhu},
  L. 2017, \aap, 597, A79

\bibitem[{{Prochaska}(2017)}]{Prochaska+17}
{Prochaska}, J.~X. 2017, Astronomy and Computing, 19, 27

\bibitem[{{Prochaska} {et~al.}(2013{\natexlab{a}}){Prochaska}, {Hennawi},
  {Lee}, {Cantalupo}, {Bovy}, {Djorgovski}, {Ellison}, {Lau}, {Martin},
  {Myers}, {Rubin}, \& {Simcoe}}]{QPQ6}
{Prochaska}, J.~X., {Hennawi}, J.~F., {Lee}, K.-G., {Cantalupo}, S., {Bovy},
  J., {Djorgovski}, S.~G., {Ellison}, S.~L., {Lau}, M.~W., {Martin}, C.~L.,
  {Myers}, A., {Rubin}, K.~H.~R., \& {Simcoe}, R.~A. 2013{\natexlab{a}}, \apj,
  776, 136

\bibitem[{{Prochaska} {et~al.}(2013{\natexlab{b}}){Prochaska}, {Hennawi}, \&
  {Simcoe}}]{QPQ5}
{Prochaska}, J.~X., {Hennawi}, J.~F., \& {Simcoe}, R.~A. 2013{\natexlab{b}},
  \apjl, 762, L19

\bibitem[{{Prochaska} {et~al.}(2014){Prochaska}, {Lau}, \& {Hennawi}}]{QPQ7}
{Prochaska}, J.~X., {Lau}, M.~W., \& {Hennawi}, J.~F. 2014, \apj, 796, 140

\bibitem[{{Prochaska} {et~al.}(2011){Prochaska}, {Weiner}, {Chen}, {Mulchaey},
  \& {Cooksey}}]{Prochaska+11}
{Prochaska}, J.~X., {Weiner}, B., {Chen}, H.-W., {Mulchaey}, J., \& {Cooksey},
  K. 2011, \apj, 740, 91

\bibitem[{{Rakic} {et~al.}(2012){Rakic}, {Schaye}, {Steidel}, \&
  {Rudie}}]{Rakic+12}
{Rakic}, O., {Schaye}, J., {Steidel}, C.~C., \& {Rudie}, G.~C. 2012, \apj, 751,
  94

\bibitem[{{Richards} {et~al.}(2002){Richards}, {Vanden Berk}, {Reichard},
  {Hall}, {Schneider}, {SubbaRao}, {Thakar}, \& {York}}]{Richards+02}
{Richards}, G.~T., {Vanden Berk}, D.~E., {Reichard}, T.~A., {Hall}, P.~B.,
  {Schneider}, D.~P., {SubbaRao}, M., {Thakar}, A.~R., \& {York}, D.~G. 2002,
  \aj, 124, 1

\bibitem[{{Rodr{\'{\i}}guez-Torres} {et~al.}(2017){Rodr{\'{\i}}guez-Torres},
  {Comparat}, {Prada}, {Yepes}, {Burtin}, {Zarrouk}, {Laurent}, {Hahn},
  {Behroozi}, {Klypin}, {Ross}, {Tojeiro}, \& {Zhao}}]{RodriguezTorres+17}
{Rodr{\'{\i}}guez-Torres}, S.~A., {Comparat}, J., {Prada}, F., {Yepes}, G.,
  {Burtin}, E., {Zarrouk}, P., {Laurent}, P., {Hahn}, C., {Behroozi}, P.,
  {Klypin}, A., {Ross}, A., {Tojeiro}, R., \& {Zhao}, G.-B. 2017, \mnras, 468,
  728

\bibitem[{{Roos} {et~al.}(2015){Roos}, {Juneau}, {Bournaud}, \&
  {Gabor}}]{Roos+15}
{Roos}, O., {Juneau}, S., {Bournaud}, F., \& {Gabor}, J.~M. 2015, \apj, 800, 19

\bibitem[{{Rubin} {et~al.}(2014){Rubin}, {Prochaska}, {Koo}, {Phillips},
  {Martin}, \& {Winstrom}}]{Rubin+14}
{Rubin}, K.~H.~R., {Prochaska}, J.~X., {Koo}, D.~C., {Phillips}, A.~C.,
  {Martin}, C.~L., \& {Winstrom}, L.~O. 2014, \apj, 794, 156

\bibitem[{{Rudie} {et~al.}(2017){Rudie}, {Newman}, \&
  {Murphy}}]{RudieNewmanMurphy17}
{Rudie}, G.~C., {Newman}, A.~B., \& {Murphy}, M.~T. 2017, \apj, 843, 98

\bibitem[{{Rupke} {et~al.}(2005){Rupke}, {Veilleux}, \& {Sanders}}]{Rupke+05}
{Rupke}, D.~S., {Veilleux}, S., \& {Sanders}, D.~B. 2005, \apjs, 160, 115

\bibitem[{{Schneider} {et~al.}(2010){Schneider}, {Richards}, {Hall}, {Strauss},
  {Anderson}, {Boroson}, {Ross}, {Shen}, {Brandt}, {Fan}, {Inada}, {Jester},
  {Knapp}, {Krawczyk}, {Thakar}, {Vanden Berk}, {Voges}, {Yanny}, {York},
  {Bahcall}, {Bizyaev}, {Blanton}, {Brewington}, {Brinkmann}, {Eisenstein},
  {Frieman}, {Fukugita}, {Gray}, {Gunn}, {Hibon}, {Ivezi{\'c}}, {Kent}, {Kron},
  {Lee}, {Lupton}, {Malanushenko}, {Malanushenko}, {Oravetz}, {Pan}, {Pier},
  {Price}, {Saxe}, {Schlegel}, {Simmons}, {Snedden}, {SubbaRao}, {Szalay}, \&
  {Weinberg}}]{Schneider+10}
{Schneider}, D.~P., {Richards}, G.~T., {Hall}, P.~B., {Strauss}, M.~A.,
  {Anderson}, S.~F., {Boroson}, T.~A., {Ross}, N.~P., {Shen}, Y., {Brandt},
  W.~N., {Fan}, X., {Inada}, N., {Jester}, S., {Knapp}, G.~R., {Krawczyk},
  C.~M., {Thakar}, A.~R., {Vanden Berk}, D.~E., {Voges}, W., {Yanny}, B.,
  {York}, D.~G., {Bahcall}, N.~A., {Bizyaev}, D., {Blanton}, M.~R.,
  {Brewington}, H., {Brinkmann}, J., {Eisenstein}, D., {Frieman}, J.~A.,
  {Fukugita}, M., {Gray}, J., {Gunn}, J.~E., {Hibon}, P., {Ivezi{\'c}}, {\v
  Z}., {Kent}, S.~M., {Kron}, R.~G., {Lee}, M.~G., {Lupton}, R.~H.,
  {Malanushenko}, E., {Malanushenko}, V., {Oravetz}, D., {Pan}, K., {Pier},
  J.~R., {Price}, III, T.~N., {Saxe}, D.~H., {Schlegel}, D.~J., {Simmons}, A.,
  {Snedden}, S.~A., {SubbaRao}, M.~U., {Szalay}, A.~S., \& {Weinberg}, D.~H.
  2010, \aj, 139, 2360

\bibitem[{{Shen} {et~al.}(2016{\natexlab{a}}){Shen}, {Brandt}, {Richards},
  {Denney}, {Greene}, {Grier}, {Ho}, {Peterson}, {Petitjean}, {Schneider},
  {Tao}, \& {Trump}}]{Shen+16}
{Shen}, Y., {Brandt}, W.~N., {Richards}, G.~T., {Denney}, K.~D., {Greene},
  J.~E., {Grier}, C.~J., {Ho}, L.~C., {Peterson}, B.~M., {Petitjean}, P.,
  {Schneider}, D.~P., {Tao}, C., \& {Trump}, J.~R. 2016{\natexlab{a}}, \apj,
  831, 7

\bibitem[{{Shen} {et~al.}(2016{\natexlab{b}}){Shen}, {Horne}, {Grier},
  {Peterson}, {Denney}, {Trump}, {Sun}, {Brandt}, {Kochanek}, {Dawson},
  {Green}, {Greene}, {Hall}, {Ho}, {Jiang}, {Kinemuchi}, {McGreer},
  {Petitjean}, {Richards}, {Schneider}, {Strauss}, {Tao}, {Wood-Vasey}, {Zu},
  {Pan}, {Bizyaev}, {Ge}, {Oravetz}, \& {Simmons}}]{Shen16}
{Shen}, Y., {Horne}, K., {Grier}, C.~J., {Peterson}, B.~M., {Denney}, K.~D.,
  {Trump}, J.~R., {Sun}, M., {Brandt}, W.~N., {Kochanek}, C.~S., {Dawson},
  K.~S., {Green}, P.~J., {Greene}, J.~E., {Hall}, P.~B., {Ho}, L.~C., {Jiang},
  L., {Kinemuchi}, K., {McGreer}, I.~D., {Petitjean}, P., {Richards}, G.~T.,
  {Schneider}, D.~P., {Strauss}, M.~A., {Tao}, C., {Wood-Vasey}, W.~M., {Zu},
  Y., {Pan}, K., {Bizyaev}, D., {Ge}, J., {Oravetz}, D., \& {Simmons}, A.
  2016{\natexlab{b}}, \apj, 818, 30

\bibitem[{{Shen} {et~al.}(2011){Shen}, {Richards}, {Strauss}, {Hall},
  {Schneider}, {Snedden}, {Bizyaev}, {Brewington}, {Malanushenko},
  {Malanushenko}, {Oravetz}, {Pan}, \& {Simmons}}]{Shen+11}
{Shen}, Y., {Richards}, G.~T., {Strauss}, M.~A., {Hall}, P.~B., {Schneider},
  D.~P., {Snedden}, S., {Bizyaev}, D., {Brewington}, H., {Malanushenko}, V.,
  {Malanushenko}, E., {Oravetz}, D., {Pan}, K., \& {Simmons}, A. 2011, \apjs,
  194, 45

\bibitem[{{Somerville} \& {Dav{\'e}}(2015)}]{SomervilleDave15}
{Somerville}, R.~S., \& {Dav{\'e}}, R. 2015, \araa, 53, 51

\bibitem[{{Steidel} {et~al.}(2010){Steidel}, {Erb}, {Shapley}, {Pettini},
  {Reddy}, {Bogosavljevi{\'c}}, {Rudie}, \& {Rakic}}]{Steidel+10}
{Steidel}, C.~C., {Erb}, D.~K., {Shapley}, A.~E., {Pettini}, M., {Reddy}, N.,
  {Bogosavljevi{\'c}}, M., {Rudie}, G.~C., \& {Rakic}, O. 2010, \apj, 717, 289

\bibitem[{{Tormen} {et~al.}(1997){Tormen}, {Bouchet}, \&
  {White}}]{TormenBouchetWhite97}
{Tormen}, G., {Bouchet}, F.~R., \& {White}, S.~D.~M. 1997, \mnras, 286, 865

\bibitem[{{Tripp} {et~al.}(2011){Tripp}, {Meiring}, {Prochaska}, {Willmer},
  {Howk}, {Werk}, {Jenkins}, {Bowen}, {Lehner}, {Sembach}, {Thom}, \&
  {Tumlinson}}]{Tripp+11}
{Tripp}, T.~M., {Meiring}, J.~D., {Prochaska}, J.~X., {Willmer}, C.~N.~A.,
  {Howk}, J.~C., {Werk}, J.~K., {Jenkins}, E.~B., {Bowen}, D.~V., {Lehner}, N.,
  {Sembach}, K.~R., {Thom}, C., \& {Tumlinson}, J. 2011, Science, 334, 952

\bibitem[{{Tumlinson} {et~al.}(2013){Tumlinson}, {Thom}, {Werk}, {Prochaska},
  {Tripp}, {Katz}, {Dav{\'e}}, {Oppenheimer}, {Meiring}, {Ford}, {O'Meara},
  {Peeples}, {Sembach}, \& {Weinberg}}]{Tumlinson+13}
{Tumlinson}, J., {Thom}, C., {Werk}, J.~K., {Prochaska}, J.~X., {Tripp}, T.~M.,
  {Katz}, N., {Dav{\'e}}, R., {Oppenheimer}, B.~D., {Meiring}, J.~D., {Ford},
  A.~B., {O'Meara}, J.~M., {Peeples}, M.~S., {Sembach}, K.~R., \& {Weinberg},
  D.~H. 2013, \apj, 777, 59

\bibitem[{{Weiner} {et~al.}(2009){Weiner}, {Coil}, {Prochaska}, {Newman},
  {Cooper}, {Bundy}, {Conselice}, {Dutton}, {Faber}, {Koo}, {Lotz}, {Rieke}, \&
  {Rubin}}]{Weiner+09}
{Weiner}, B.~J., {Coil}, A.~L., {Prochaska}, J.~X., {Newman}, J.~A., {Cooper},
  M.~C., {Bundy}, K., {Conselice}, C.~J., {Dutton}, A.~A., {Faber}, S.~M.,
  {Koo}, D.~C., {Lotz}, J.~M., {Rieke}, G.~H., \& {Rubin}, K.~H.~R. 2009, \apj,
  692, 187

\bibitem[{{Wild} {et~al.}(2008){Wild}, {Kauffmann}, {White}, {York}, {Lehnert},
  {Heckman}, {Hall}, {Khare}, {Lundgren}, {Schneider}, \& {vanden
  Berk}}]{Wild+08}
{Wild}, V., {Kauffmann}, G., {White}, S., {York}, D., {Lehnert}, M., {Heckman},
  T., {Hall}, P.~B., {Khare}, P., {Lundgren}, B., {Schneider}, D.~P., \&
  {vanden Berk}, D. 2008, \mnras, 388, 227

\bibitem[{{Zahedy} {et~al.}(2016){Zahedy}, {Chen}, {Rauch}, {Wilson}, \&
  {Zabludoff}}]{Zahedy+16}
{Zahedy}, F.~S., {Chen}, H.-W., {Rauch}, M., {Wilson}, M.~L., \& {Zabludoff},
  A. 2016, \mnras, 458, 2423

\end{thebibliography}

\end{document}