### 2018-02-13 QPQ9/Submit/response_2.txt

Dear referee,

We thank you very much for careful reading of our manuscript the second
time. Your constructive criticism has been very helpful.

>=== Major concern 1: failure to capture velocity spread of the absorption
>and misuse of the central limit theorem
>
>In response to one of my concerns, the authors now motivate Gaussian
>fitting to the mean absorption profile using the central limit theorem.
>This is a fundamental misuse of the central limit theorem. The central
>limit theorem implies (under fairly loose assumptions) that the sums of
>random variables will tend towards a normal distribution. This means, for
>example, that the measured flux in a particular pixel of the stack will be
>normally >distributed. But it *does not* imply that the mean line profiles
>themsleves will be Gaussian. While there are conditions under which mean
>line profiles will converge to a Gaussian, these are significantly more
>demanding conditions that typically do not apply to astrophysical
>measurements and objects. For example, the mean stack of point sources in
>ground based images will not converge to a Gaussian because of the large
>fraction of flux contained in wings of the PSF. Similarly, mean line
>profiles in stacks are often far from Gaussian such as: both broad and
>narrow quasar emission lines (Schneider+2010, Shen+2016b), galaxy emission
>lines (Cicone+2016), and associated quasar absorption lines which are
>distinctly non-Gaussian as shown in Figure 9 of this paper.
>
>Previous QPQ papers (Prochaska & Hennawi 2009, Lau+2016) and those from
>other authors (e.g. Tripp+2011, Johnson+2015, Rudie+2017) have shown that
>AGN and quasar hosts often exhibit absorbing gas at "extreme" velocities
>of 500-1000 km/s. Indeed, these extreme velocities are discussed in the
>introduction. However, it is clear that the extreme velocities are not
>being captured by the measured Gaussian fit.
>
>This concern is raised not only by the previous works but also by careful
>examination of Figure 2. In particular the pseudo-continuum level in the
>mean stack is systematically lower than in the median stack for all three
>ions considered. This indicates that there is additional absorption at
>large velocities which is not being captured by the analysis in the paper
>but is instead being erroneously folded into the pseudo-continuum
>estimate.
>
>This is a major issue with the adopted methodology that must be addressed
>with careful thought, new experiments/testing, and detailed discussion.

We thank the referee for pointing out the misuse of central limit theorem!
It applies to the fluxes at each pixel. The absorption in the stack will
follow the distribution from which it was sampled, by the law of large
numbers. We present a more careful argument for fitting one Gaussian
component and one constant component, and removed the text relying on the
central limit theorem.

The distribution of absorbers unassociated with the foreground quasars is
expected to be uniform and random, and depress the continuum level in the
stack uniformly. This justifies our fitting of a pseudocontinuum, and is
not the quantity of interest. The questions to be addressed is, whether
one component is a good description of the dispersion in the average
absorption, and if so, whether the dispersion in this component is well
captured by a Gaussian. We tried replacing a constant pseudocontinuum with
a second Gaussian absorption component in our modeling of the CII stack.
We found this second Gaussian component to be weak and broad. The amplitude
is 0.03, compared to our original constant component of 0.97. The standard
deviation is 3600 km/s, whose span is wider than the entire velocity window
for stacking. This weak and very broad component also has little effect on
the centroid and width of the narrower Gaussian component. We therefore see
no convincing evidence for a broader component hidden in the depressed
continuum. We then calculated the dispersion in a +/-1300 km/s window
surrounding the absorption centroidusing the standard deviation formula. We
have this large velocity window to encompass most of the associated
absorption. We found the dispersion measured this way is essentially the
same as that found by fitting a Gaussian. This also leans support to our
approach of fitting two Gaussians to the CIV and MgII doublet, where we
cannot directly apply the standard deviation formula.

In our revised manuscript, we acknowledge the possibility our Gaussian
model does not capture all the powers at extreme velocities. However we
also note that occasional large kinematic offsets do not make a strong case
against gravitational motions. Firstly, quasars occasionally inhabit
extremely overdense environments, such as protoclusters. Secondly, the
probability of intercepting a random absorber is boosted by large scale
clustering. Hence occasional extreme velocities would not necessitate
outflows. Our preferred kinematic measure is thus the dispersion in the
average absorption. We added the above text to the revised manuscript.

The uniformly depressed continuum results from absorbers unrelated to the
foreground quasars, which is expected to have a covering fraction less than
50%. For median stack, a pixel is not affected unless there are more than
50% of quasar pairs with absorption at its velocity. The pseudo-continuum
of the mean stack being lower than the median stack is not a reflection of
associated absorption at large velocities.

>=== Major concern 2: insufficient detail in how the doublet fitting is
>performed.
>
>I thank the authors for specifying the allowed range of doublet ratio.
>However, additional information is still needed on the doublet fitting,
>particularly regarding the pseudo continuum.
>
>The doublet ratio is degenerate with centroid in some cases, particularly
>for CIV and this possible degeneracy is not discussed.
>
>Depending on how the pseudo continuum fitting is performed, it could
>potentially suppress any broad wings in the mean absorption profile. The
>authors have most likely taken steps to prevent this from happening. But
>if so, these steps are note specified.f
>
>Provide additional details addressing both of these points.

We tried fitting the continuum separately in (-3000,-1300) km/s and (+1300,
+3000) km/s instead of multiplying a constant to a Gaussian. We found the
continuum determined this way is essentially the same.

We acknowledge the degeneracy between the centroid and the doublet ratio
for CIV. And hence, we state that the results for the doublets are
presented for consistency check, but the analysis focuses on CII. We
emphasized this point in the revised manuscript. For CIV, a doublet ratio
closer 2:1 is preferred. We found a doublet ratio of 2:1 captures the
absorption at the central few pixels better than a doublet ratio of 1:1
does.

>=== Main concern 3: mis-representation of papers in the literature
>
>While most of these have been corrected, one remains and a new one has
>been added. In particular, both Tripp+2011 and Rudie+2017 found absorbing
>gas spread over ~1000 km/s which is comparable to the velocity spreads
>from the QPQ series that are discussed as extreme in the introduction. The
>main conclusions of Tripp+2011 and Rudie+2017 focus on these large
>velocity spreads and the current description in the introduction is
>therefore misleading.

>We thank the referee for scrutinizing our representation of the
>literature. We dropped the words that claim these absorption systems have
>most of the gas within several hundred km/s.

>=== In addition:
>
>A) To more convincingly demonstrate that the velocity shift in the
>measured mean absorption profile is real the authors should:
>
>1) Remove possible contamination from absorption systems associated with
>the background quasar. The paper stacks the background spectra for pairs
>in which the redshift difference between the foreground and background
>quasar is >3000 km/s in order to prevent contamination by physical quasar
>pairs in the presence of redshift uncertainty. However, quasars also
>exhibit narrow associated absorption systems at large velocities of up to
>~12,000 km/s (Wild+2008) which could also artificially introduce an
>asymmetry into the experiment. The authors can easily demonstrate that
>this is no tthe case by repeating the stacking procedure on quasars pairs
>with a more restrictive range of allowed velocity separation of >12,000
>km/s.
>
>2) Demonstrating that CII* cannot bias the measurement of the CII velocity
>centroid. Previous QPQ papers have identified CII* absorption in the CGM
>of quasars. This absorption feature is blue-shifted from the ground state
>transition by 260 km/s which is comparable to the measured offset. With
>data from Lau+2016, the authors can easily demonstrate that contamination
>from CII* is not responsible for the observed shift.

We identified and excluded all spectra with BAL and mini-BAL systems.
Possible contamination from narrow associated absorbers of background
quasars should result in a tilt in the pseudo-continuum. Associated
absorbers of background quasars will not prefer the systemic velocity of
the foreground quasars, and hence will not affect the average absorption
measured for the foreground quasars. The median pair velocity separation
of the CII sample is 11140 km/s. Dividing the sample ino two bins of pair
velocity separation does not result in meaningful comparison, due to the
smaller number of pairs in each bin. Hwever we see no evidence for a
difference in the centroid's redshift of the two bins.

We examined all the nine CII-bearing subsystems in QPQ8 where CII* is not
blended with CII. Among the above nine subsystems, only one, J1420
subsystem F, has CII* equivalent width comparable to CII. Hence,
contamination from CII* is insignifcant and could not be responsible for
the observed shift. We thank the referee for suggesting us to add this to
the text.

>B) To demonstrate that the quasar variability effect can plausibly
>produce the observed shift from Hubble flow alone the authors need to
>calculate the distance to which the quasars can substantially contribute
>to the UV background and convert this distance to a line-of-sight
>velocity assuming Hubble flow. Most of the work for this has already been
>done for the calculation of g_UV for Figure 1 and Table 1 so only a few
>additional steps are needed. Implementing this calculation in the Monte
>Carlo approach in Section 3.3 would be particularly convincing.

In our test of implementing an arbitrary quasar opening angle in our Monte
Carlo simulations to reproduce the observed shift, we remove all absorbers
at CGM distances that fall within the opening angle to mimic UV
enhancement. We did not remove absorbers at larger distances/Hubble flow
velocities as the effect of quasar variability is expected to be weaker.
However, for the stack of mock spectra fro Monte Carlo simulations that we
present in Figure 5, we did not include this quasar variability effect but
merely shifted the absorption centroid to match with the observations. The
goal of Figure 5 is to show that the observed velocity width can be
reproduced without outflows. Introducing quasar variability would
introduce more free parameters and the figure may be less informative.

>C) To clarify the discussion the authors should:
>1) Rephrase or expand the sentence/paragraph: "As a test of
>reasonableness, we calculate that, for a quasar opening angle of 120...".
>It is not clear what calculation was performed and how it represents a
>test of reasonableness.
>2) Remove the sentence "Light from the background quasar will travel
>smaller  distance to reach the gas behind the foreground quasar than the
>gas in front of it." It is repetitive and the subsequent sentence
>explains this physical picture more clearly.
>3) The last sentence of the paper is misleading because the observed
>velocity shift cannot be explained by gravitational motion.

We expanded the text on the test of an arbitrary, reasonable quasar
opening angle. In this test, we removed all CGM absorbers that fall
within the quasar opening angle in our Monte Carlo simulations. We found
that for an opening angle of 120 degrees, the observed centroid may be
reproduced while the intrinsic dispersion is still wthin the modeling
error.
We removed the sentence "Light from the background quasar will travel..."
as the referee suggested. This indeed helps clear up the test.
We rewrote the last sentence to separately provide the explanation for
the large velocity widths and the explanation for the average shift.

>Finally, I would like to thank the authors for going the extra mile and
>researching color choices that improve figure legibility for color blind
>readers. However, the stated change does not appear to have been
>implemented in Figure 6. Since the work has already been done,
>implementing the color change would be an easy correction.

We examined the cartoon figures again. A lighter shade of blue instead
of navy had been used for the CIV ions. It seems the poor legibility also
came from the figures being too compact. We enlarged the figures in the
revised manuscript. We are very happy to use another color scheme should
the referee or editor have better suggestion.

Marie Wingyee Lau