### 2018-02-27 QPQ9/Submit/response_3.txt
response to third referee report

Dear referee,

We thank you very much for the careful reading and constructive criticism
that will help toward the eventual acceptance for publication.

>Remaining major question 1:
>
>In the response to the second referee report and in the caption of
>Figure 5, the authors state that the Monte Carlo simulation absorption
>is shifted to match the centroid of the absorption.
>Am I correct in understanding that this is done to force a match
>between the observed absorption shift and the Monte Carlo model? If so,
>this would seem to invalidate the claim by the paper that the observed
>velocity shift can be explained by anisotropic and/or intermittent quasar
>radiation combined with the Hubble flow. Why can't the shift be
>reproduced by the Monte Carlo model which includes all of these effects?
>If the model cannot reproduce the shift, then some of the major
>conclusions of the paper, including the last sentence of the abstract,
>are not supported. I am sufficiently puzzled by this that I request the
>opportunity to review the paper one more time before acceptance. Please
>clarify and explain the need for an artificial velocity shift in order
>for the Monte Carlo model to match the observed stack.

We thank the referee very much for pushing us to examine our Monte Carlo
simulation outputs. Asymmetric distribution of Hubble velocities alone
cannot produce the observed redshifted centroid. We made a mistake in
our claim in the last response, as we mistook the simulation outputs
that incorporated outflows to be those without outflows. We find that,
with asymmetric distribution in Hubble velocities alone, the centroid
may at most be shifted ~80 km/s. Hence, the observed putative asymmetry
suggests the presence of an outflowing component! We crudely estimate
that, the observed dispersion and centroid may be reproduced with a
quasar opening angle of 120 degrees and radial velocities of 500 km/s.
We added text that justifies ionizing absorbers at Mpc distances from
the quasar. The argument comes from the observed anisotropic clustering
of absorbers around quasars.

We added text about outflows. In addition, we also restructured the text
somewhat to first discuss the observed dispersion and then discuss the
observed centroid. We put a confidence limit on the observed centroid
of the average absorption from CII at 232+/-98 km/s, with large intrinsic
scatter. Hence, we consider it worthwhile to also separately analyze the
observed disperson.

>Remaining major question 2:
>
>In the Monte Carlo model, the authors adopt a sigma_rms scatter of 246
>km/s based on the expected rms velocity for quasar host halos. But the
>absorption systems being considered in the model are in the Hubble flow
>at significant cosmological distance, not in the quasar host halo. Why
>then is a sigma_rms from massive halos appropriate? Does the adoption of
>a different sigma_rms change the results?

We thank the referee for this suggestion, that helps strengthen the
results. We try adopting a sigma_rms scatter more more typical of coeval
Lyman break galaxies, for absorbers at cosmological distances. We find
that the dispersion in the model is again consistent with the
observation. Relevant text has been added.

>Other points for improvement:

>1) Explicitly state the flux level of the psuedo-continuum by adding it
>to Table 2
>
>2) Similarly, the paper must explicitly give the width of the second
>Guassian component and equivalent width discussed on page 6. This means
>providing actual numbers with uncertainties in the text rather than
>qualitative assertions.
>
>3) The discussion of a bias from CII* is not convincing as is (though I
>am certain that the claim of small bias is correct). In order to make the
>given argument convincing, the authors need to give quantitative
>constraints on the mean CII to CII* ratio and then translate this ratio
>into an estimated bias. Again, the paper must include specific,
>quantitative information on the estimated level of this bias rather than
>simply assert that it is insignificant.
>
>4) On the color scheme in Figure 6: Anything that does not involve red
>and green would be a significant improvement. This figure could be made
>in greyscale, coolmap, heatmap, or any number of other options.
>Alternative, a color scheme could be dropped entirely by plotting the 1,
>2, and 3 sigma regions in solid, dashed, and dotted line respectively.
>Or different hatching schemes could be used for each reason.

We have added the pseudo-continuum levels to Table 2.

We also added text to quantify the width, amplitude, and uncertainties in
the wider and waeker Gaussian component.

The Mean CII to CII* ratio is 0.2, and we estimate that CII* may at most
add a ~40 km/s positive bias.

In our response to the last report, we overlooked the referee's request
for changing the color scheme of Figure 6, and we had only done that for
the two cartoon figures. We apologize for that, and we have now changed
the color scheme of Figure 6.

Marie Wingyee Lau