### 2018-04-27 QPQ9/Submit/response_4.txt

Dear referee,

It's been very helpful to have these exchanges with you. Thank you
very much for your professional service. I have added your latest
suggestions.

>1) The paper now states that a 500 km/s outflow can reproduce the
>observed velocity shift based on "crude" estimates. Please give a bit
>more detail on how this estimate was obtained and state whether it is
>degenerate with e.g. the quasar opening angle or episodic lifetime.

We updated the text to be somewhat more quantitative and rigorous.
Anistropic emission is degenerate with intermittent emission.
First we consider the case where the quasar lifetime is infinite, and
the observed asymmetry is only produced by anisotropic emission. Then,
the radial outflow speed needs to be ~450 km/s and the opening
angle needs to be ~180 degrees, so that the observed intrinsic
dispersion and absorption centroid can be produced.
Second we consider the case where the quasar is isotropic, and the
observed asymmetry is only produced by finite lifetime. We further make
the assumption the quasar luminosity is constant during its lifetime.
Then, the radial outflow speed needs to be at least ~420 km/s and the
quasar must have shined for ~0.4e+6 years, to reproduce the observation.
For the anisotropy-only case, the opening angle
deduced is rather large compared to literature findings (e.g.
Borisova+16). For the intermittence-only case, if the quasars are on
averaged observed near the middle of the episode, the lifetime
deduced is somewhat small compared to literature findings. We thus
speculate that the asymmetry is the result of a combination of anisotropic
and intermittent emission. To better illustrate the second case, we also
updated the cartoon figure on finite lifetime to show a scenario where the
emission is isotropic.

>2) It would be useful to compare this outflow rate with other works and
>discuss possible origins of the putative outflows (e.g. supernovae vs.
>AGN). For example, a comparison with Turner+2014, 2017 would be
>enlightening. In particulary, Turner+2017 examined mean absorption
>optical depths around Lyman break galaxies and compared with simulations.
>For lower ionization state gas (HI, SiIV, and CIV in this case), the
>favored conclusion is that clustering of the absorption in velocity space
>is most consistent with gas that is *inflowing* on average. This is in
>apparent contrast with the new conclusions of the paper. This difference
>may be related to the masses of the systems being studied or the presence
>of the quasar due to quasar driven outflows (e.g. Green+2012) and/or
>supernovae driven outflows from starburst activity that is correlated
>with the presence of quasars (e.g. Barthel+2017).

We thank the referee for the literature suggestion, and the discussion
suggestion. We have added relevant text.

>1) It would be useful to add minor ticks to Figure 3, particularly on the
>x-axis.

We added minor ticks. We thank the referee's attentiveness to details.

>2) The numbering on the x-axis of Figure 4 is difficult to read
>currently.

We decreased the number of tick labels. We thank the referee for their
attentiveness.

Marie Wingyee Lau